### 2012-05-04 paper/paper.tex
Paper templates

% This is "sig-alternate.tex" V1.9 April 2009
% This file should be compiled with V2.4 of "sig-alternate.cls" April 2009
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.4 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.4) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V1.9 - April 2009

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
and don't forget to end the environment with
{figure*}, not {figure}!

\begin{figure*}
\centering
\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\begin{figure}
\centering
\psfig{file=rosette.ps, height=1in, width=1in,}
\caption{A sample black and white graphic (.ps format) that has
been resized with the \texttt{psfig} command.}
\vskip -6pt
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The sig-alternate.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2012-05-04 paper/paper.tex
Paper stuff

@@
-% This is "sig-alternate.tex" V1.9 April 2009
-% This file should be compiled with V2.4 of "sig-alternate.cls" April 2009
-%
-% This example file demonstrates the use of the 'sig-alternate.cls'
-% V2.4 LaTeX2e document class file. It is for those submitting
-% articles to ACM Conference Proceedings WHO DO NOT WISH TO
-% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
-% The 'sig-alternate.cls' file will produce a similar-looking,
-% albeit, 'tighter' paper resulting in, invariably, fewer pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.4) produces:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which
-% DOES NOT produce 1) thru' 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within
-% the source .tex file, over both the CopyrightYear
-% (defaulted to 200X) and the ACM Copyright Data
-% (defaulted to X-XXXXX-XX-X/XX/XX).
-% e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use
-% the .bib file (from which the .bbl file % is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission, you *NEED* to 'insert'
-% your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% ================= IF YOU HAVE QUESTIONS =======================
-% Questions regarding the SIGS styles, SIGS policies and
-% procedures, Conferences etc. should be sent to
-% Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to
-% Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V1.9 - April 2009
-
 \documentclass{sig-alternate}
@@
 \begin{document}
-%
-% --- Author Metadata here ---
-\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-% --- End of Author Metadata ---
 
-\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
-Format\titlenote{(Produces the permission block, and
-copyright information). For use with
-SIG-ALTERNATE.CLS. Supported by ACM.}}
-\subtitle{[Extended Abstract]
-\titlenote{A full version of this paper is available as
-\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-\LaTeX$2_\epsilon$\ and BibTeX} at
-\texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
+\CopyrightYear{2012}
 
-\numberofauthors{8} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
+\title{Motmot: A Secure, Distributed Chat Service}
+
+\numberofauthors{5}
 %
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
+\alignauthor
+Carl Jackson \\
+  \affaddr{Harvard University} \\
+  \email{cjackson@college.harvard.edu}
 %
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
+\alignauthor
+Max Wang \\
+  \affaddr{Harvard University} \\
+  \email{max.wang@college.harvard.edu}
 %
-% 1st. author
+\and
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
-% 2nd. author
+EJ Bensing \\
+  \affaddr{Harvard University} \\
+  \email{ebensing@college.harvard.edu}
+%
 \alignauthor
-G.K.M. Tobin\titlenote{The secretary disavows
-any knowledge of this author's actions.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{P.O. Box 1212}\\
-       \affaddr{Dublin, Ohio 43017-6221}\\
-       \email{webmaster@marysville-ohio.com}
-% 3rd. author
-\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-one who did all the really hard work.}\\
-       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-       \affaddr{Hekla, Iceland}\\
-       \email{larst@affiliation.org}
-\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-\alignauthor Lawrence P. Leipuner\\
-       \affaddr{Brookhaven Laboratories}\\
-       \affaddr{Brookhaven National Lab}\\
-       \affaddr{P.O. Box 5000}\\
-       \email{lleipuner@researchlabs.org}
-% 5th. author
-\alignauthor Sean Fogarty\\
-       \affaddr{NASA Ames Research Center}\\
-       \affaddr{Moffett Field}\\
-       \affaddr{California 94035}\\
-       \email{fogartys@amesres.org}
-% 6th. author
-\alignauthor Charles Palmer\\
-       \affaddr{Palmer Research Laboratories}\\
-       \affaddr{8600 Datapoint Drive}\\
-       \affaddr{San Antonio, Texas 78229}\\
-       \email{cpalmer@prl.com}
+Jeff Atwood \\
+  \affaddr{Harvard University} \\
+  \email{jatwood@college.harvard.edu}
+%
+\alignauthor
+Julie Hansbrough \\
+  \affaddr{Harvard University} \\
+  \email{jhansbrough@college.harvard.edu}
 }
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{30 July 1999}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
 
 \maketitle
+
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms,
-somewhat loosely, to the formatting guidelines for
-ACM SIG Proceedings. It is an {\em alternate} style which produces
-a {\em tighter-looking} paper and was designed in response to
-concerns expressed, by authors, over page-budgets.
-It complements the document \textit{Author's (Alternate) Guide to
-Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
-This source file has been written with the intention of being
-compiled under \LaTeX$2_\epsilon$\ and BibTeX.
 
-The developers have tried to include every imaginable sort
-of ``bells and whistles", such as a subtitle, footnotes on
-title, subtitle and authors, as well as in the text, and
-every optional component (e.g. Acknowledgments, Additional
-Authors, Appendices), not to mention examples of
-equations, theorems, tables and figures.
+We should write an abstract here.
 
-To make best use of this sample document, run it through \LaTeX\
-and BibTeX, and compare this source code with the printed
-output produced by the dvi file. A compiled PDF version
-is available on the web page to help you with the
-`look and feel'.
 \end{abstract}
 
-% A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+\section{Introduction}
 
-\terms{Theory}
-
-\keywords{ACM proceedings, \LaTeX, text tagging}
-
-\section{Introduction}
-The \textit{proceedings} are the records of a conference.
-ACM seeks to give these conference by-products a uniform,
-high-quality appearance.  To do this, ACM has some rigid
-requirements for the format of the proceedings documents: there
-is a specified format (balanced  double columns), a specified
-set of fonts (Arial or Helvetica and Times Roman) in
-certain specified sizes (for instance, 9 point for body copy),
-a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
-the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
-and (1.9 cm [.75"]) left and right; specified column width
-(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).
-
-The good news is, with only a handful of manual
-settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
-and {\texttt{\char'134 alignauthor}} commands, you have
-already used; another, {\texttt{\char'134 balancecolumns}}, will
-be used in your very last run of \LaTeX\ to ensure
-balanced column heights on the last page.}, the \LaTeX\ document
-class file handles all of this for you.
-
-The remainder of this document is concerned with showing, in
-the context of an ``actual'' document, the \LaTeX\ commands
-specifically available for denoting the structure of a
-proceedings paper, rather than with giving rigorous descriptions
-or explanations of such commands.
-
-\section{The {\secit Body} of The Paper}
-Typically, the body of a paper is organized
-into a hierarchical structure, with numbered or unnumbered
-headings for sections, subsections, sub-subsections, and even
-smaller sections.  The command \texttt{{\char'134}section} that
-precedes this paragraph is part of such a
-hierarchy.\footnote{This is the second footnote.  It
-starts a series of three footnotes that add nothing
-informational, but just give an idea of how footnotes work
-and look. It is a wordy one, just so you see
-how a longish one plays out.} \LaTeX\ handles the numbering
-and placement of these headings for you, when you use
-the appropriate heading commands around the titles
-of the headings.  If you want a sub-subsection or
-smaller part to be unnumbered in your output, simply append an
-asterisk to the command name.  Examples of both
-numbered and unnumbered headings will appear throughout the
-balance of this sample document.
-
-Because the entire article is contained in
-the \textbf{document} environment, you can indicate the
-start of a new paragraph with a blank line in your
-input file; that is why this sentence forms a separate paragraph.
-
-\subsection{Type Changes and {\subsecit Special} Characters}
-We have already seen several typeface changes in this sample.  You
-can indicate italicized words or phrases in your text with
-the command \texttt{{\char'134}textit}; emboldening with the
-command \texttt{{\char'134}textbf}
-and typewriter-style (for instance, for computer code) with
-\texttt{{\char'134}texttt}.  But remember, you do not
-have to indicate typestyle changes when such changes are
-part of the \textit{structural} elements of your
-article; for instance, the heading of this subsection will
-be in a sans serif\footnote{A third footnote, here.
-Let's make this a rather short one to
-see how it looks.} typeface, but that is handled by the
-document class file. Take care with the use
-of\footnote{A fourth, and last, footnote.}
-the curly braces in typeface changes; they mark
-the beginning and end of
-the text that is to be in the different typeface.
-
-You can use whatever symbols, accented characters, or
-non-English characters you need anywhere in your document;
-you can find a complete list of what is
-available in the \textit{\LaTeX\
-User's Guide}\cite{Lamport:LaTeX}.
-
-\subsection{Math Equations}
-You may want to display math equations in three distinct styles:
-inline, numbered or non-numbered display.  Each of
-the three are discussed in the next sections.
-
-\subsubsection{Inline (In-text) Equations}
-A formula that appears in the running text is called an
-inline or in-text formula.  It is produced by the
-\textbf{math} environment, which can be
-invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
-construction or with the short form \texttt{\$. . .\$}. You
-can use any of the symbols and structures,
-from $\alpha$ to $\omega$, available in
-\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
-few examples of in-text equations in context. Notice how
-this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
-set here in in-line math style, looks slightly different when
-set in display style.  (See next section).
-
-\subsubsection{Display Equations}
-A numbered display equation -- one set off by vertical space
-from the text and centered horizontally -- is produced
-by the \textbf{equation} environment. An unnumbered display
-equation is produced by the \textbf{displaymath} environment.
-
-Again, in either environment, you can use any of the symbols
-and structures available in \LaTeX; this section will just
-give a couple of examples of display equations in context.
-First, consider the equation, shown as an inline equation above:
-\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
-Notice how it is formatted somewhat differently in
-the \textbf{displaymath}
-environment.  Now, we'll enter an unnumbered equation:
-\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
-and follow it with another numbered equation:
-\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
-just to demonstrate \LaTeX's able handling of numbering.
-
-\subsection{Citations}
-Citations to articles \cite{bowman:reasoning,
-clark:pct, braams:babel, herlihy:methodology},
-conference proceedings \cite{clark:pct} or
-books \cite{salas:calculus, Lamport:LaTeX} listed
-in the Bibliography section of your
-article will occur throughout the text of your article.
-You should use BibTeX to automatically produce this bibliography;
-you simply need to insert one of several citation commands with
-a key of the item cited in the proper location in
-the \texttt{.tex} file \cite{Lamport:LaTeX}.
-The key is a short reference you invent to uniquely
-identify each work; in this sample document, the key is
-the first author's surname and a
-word from the title.  This identifying key is included
-with each item in the \texttt{.bib} file for your article.
-
-The details of the construction of the \texttt{.bib} file
-are beyond the scope of this sample document, but more
-information can be found in the \textit{Author's Guide},
-and exhaustive details in the \textit{\LaTeX\ User's
-Guide}\cite{Lamport:LaTeX}.
-
-This article shows only the plainest form
-of the citation command, using \texttt{{\char'134}cite}.
-This is what is stipulated in the SIGS style specifications.
-No other citation format is endorsed or supported.
-
-\subsection{Tables}
-Because tables cannot be split across pages, the best
-placement for them is typically the top of the page
-nearest their initial cite.  To
-ensure this proper ``floating'' placement of tables, use the
-environment \textbf{table} to enclose the table's contents and
-the table caption.  The contents of the table itself must go
-in the \textbf{tabular} environment, to
-be aligned properly in rows and columns, with the desired
-horizontal and vertical rules.  Again, detailed instructions
-on \textbf{tabular} material
-is found in the \textit{\LaTeX\ User's Guide}.
-
-Immediately following this sentence is the point at which
-Table 1 is included in the input file; compare the
-placement of the table here with the table in the printed
-dvi output of this document.
-
-\begin{table}
-\centering
-\caption{Frequency of Special Characters}
-\begin{tabular}{|c|c|l|} \hline
-Non-English or Math&Frequency&Comments\\ \hline
-\O & 1 in 1,000& For Swedish names\\ \hline
-$\pi$ & 1 in 5& Common in math\\ \hline
-\$ & 4 in 5 & Used in business\\ \hline
-$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-\hline\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of
-the page's live area, use the environment
-\textbf{table*} to enclose the table's contents and
-the table caption.  As with a single-column table, this wide
-table will ``float" to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table 2 is included in the input file; again, it is
-instructive to compare the placement of the
-table here with the table in the printed dvi
-output of this document.
-
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-% end the environment with {table*}, NOTE not {table}!
-
-\subsection{Figures}
-Like tables, figures cannot be split across pages; the
-best placement for them
-is typically the top or the bottom of the page nearest
-their initial cite.  To ensure this proper ``floating'' placement
-of figures, use the environment
-\textbf{figure} to enclose the figure and its caption.
-
-This sample document contains examples of \textbf{.eps}
-and \textbf{.ps} files to be displayable with \LaTeX.  More
-details on each of these is found in the \textit{Author's Guide}.
-
-\begin{figure}
-\centering
-\epsfig{file=fly.eps}
-\caption{A sample black and white graphic (.eps format).}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=fly.eps, height=1in, width=1in}
-\caption{A sample black and white graphic (.eps format)
-that has been resized with the \texttt{epsfig} command.}
-\end{figure}
-
-As was the case with tables, you may want a figure
-that spans two columns.  To do this, and still to
-ensure proper ``floating'' placement of tables, use the environment
-\textbf{figure*} to enclose the figure and its caption.
-and don't forget to end the environment with
-{figure*}, not {figure}!
-
-\begin{figure*}
-\centering
-\epsfig{file=flies.eps}
-\caption{A sample black and white graphic (.eps format)
-that needs to span two columns of text.}
-\end{figure*}
-
-Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
-used; use
-the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
-commands as appropriate for the different file types.
-
-\begin{figure}
-\centering
-\psfig{file=rosette.ps, height=1in, width=1in,}
-\caption{A sample black and white graphic (.ps format) that has
-been resized with the \texttt{psfig} command.}
-\vskip -6pt
-\end{figure}
-
-\subsection{Theorem-like Constructs}
-Other common constructs that may occur in your article are
-the forms for logical constructs like theorems, axioms,
-corollaries and proofs.  There are
-two forms, one produced by the
-command \texttt{{\char'134}newtheorem} and the
-other by the command \texttt{{\char'134}newdef}; perhaps
-the clearest and easiest way to distinguish them is
-to compare the two in the output of this sample document:
-
-This uses the \textbf{theorem} environment, created by
-the\linebreak\texttt{{\char'134}newtheorem} command:
-\newtheorem{theorem}{Theorem}
-\begin{theorem}
-Let $f$ be continuous on $[a,b]$.  If $G$ is
-an antiderivative for $f$ on $[a,b]$, then
-\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
-\end{theorem}
-
-The other uses the \textbf{definition} environment, created
-by the \texttt{{\char'134}newdef} command:
-\newdef{definition}{Definition}
-\begin{definition}
-If $z$ is irrational, then by $e^z$ we mean the
-unique number which has
-logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
-\end{definition}
-
-Two lists of constructs that use one of these
-forms is given in the
-\textit{Author's  Guidelines}.
-
-There is one other similar construct environment, which is
-already set up
-for you; i.e. you must \textit{not} use
-a \texttt{{\char'134}newdef} command to
-create it: the \textbf{proof} environment.  Here
-is a example of its use:
-\begin{proof}
-Suppose on the contrary there exists a real number $L$ such that
-\begin{displaymath}
-\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-\end{displaymath}
-Then
-\begin{displaymath}
-l=\lim_{x\rightarrow c} f(x)
-= \lim_{x\rightarrow c}
-\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
-= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-\frac{f(x)}{g(x)} = 0\cdot L = 0,
-\end{displaymath}
-which contradicts our assumption that $l\neq 0$.
-\end{proof}
-
-Complete rules about using these environments and using the
-two different creation commands are in the
-\textit{Author's Guide}; please consult it for more
-detailed instructions.  If you need to use another construct,
-not listed therein, which you want to have the same
-formatting as the Theorem
-or the Definition\cite{salas:calculus} shown above,
-use the \texttt{{\char'134}newtheorem} or the
-\texttt{{\char'134}newdef} command,
-respectively, to create it.
-
-\subsection*{A {\secit Caveat} for the \TeX\ Expert}
-Because you have just been given permission to
-use the \texttt{{\char'134}newdef} command to create a
-new form, you might think you can
-use \TeX's \texttt{{\char'134}def} to create a
-new command: \textit{Please refrain from doing this!}
-Remember that your \LaTeX\ source code is primarily intended
-to create camera-ready copy, but may be converted
-to other forms -- e.g. HTML. If you inadvertently omit
-some or all of the \texttt{{\char'134}def}s recompilation will
-be, to say the least, problematic.
+Wooooo.
 
 \section{Conclusions}
-This paragraph will end the body of this sample document.
-Remember that you might still have Acknowledgments or
-Appendices; brief samples of these
-follow.  There is still the Bibliography to deal with; and
-we will make a disclaimer about that here: with the exception
-of the reference to the \LaTeX\ book, the citations in
-this paper are to articles which have nothing to
-do with the present subject and are used as
-examples only.
-%\end{document}  % This is where a 'short' article might terminate
 
-%ACKNOWLEDGMENTS are optional
+Wheeee.
+
 \section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
 
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-%\balancecolumns
-\appendix
-%Appendix A
-\section{Headings in Appendices}
-The rules about hierarchical headings discussed above for
-the body of the article are different in the appendices.
-In the \textbf{appendix} environment, the command
-\textbf{section} is used to
-indicate the start of each Appendix, with alphabetic order
-designation (i.e. the first is A, the second B, etc.) and
-a title (if you include one).  So, if you need
-hierarchical structure
-\textit{within} an Appendix, start with \textbf{subsection} as the
-highest level. Here is an outline of the body of this
-document in Appendix-appropriate form:
-\subsection{Introduction}
-\subsection{The Body of the Paper}
-\subsubsection{Type Changes and  Special Characters}
-\subsubsection{Math Equations}
-\paragraph{Inline (In-text) Equations}
-\paragraph{Display Equations}
-\subsubsection{Citations}
-\subsubsection{Tables}
-\subsubsection{Figures}
-\subsubsection{Theorem-like Constructs}
-\subsubsection*{A Caveat for the \TeX\ Expert}
-\subsection{Conclusions}
-\subsection{Acknowledgments}
-\subsection{Additional Authors}
-This section is inserted by \LaTeX; you do not insert it.
-You just add the names and information in the
-\texttt{{\char'134}additionalauthors} command at the start
-of the document.
-\subsection{References}
-Generated by bibtex from your ~.bib file.  Run latex,
-then bibtex, then latex twice (to resolve references)
-to create the ~.bbl file.  Insert that ~.bbl file into
-the .tex source file and comment out
-the command \texttt{{\char'134}thebibliography}.
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-\section{More Help for the Hardy}
-The sig-alternate.cls file itself is chock-full of succinct
-and helpful comments.  If you consider yourself a moderately
-experienced to expert user of \LaTeX, you may find reading
-it useful but please remember not to change it.
-%\balancecolumns % GM June 2007
-% That's all folks!
+\begin{thebibliography}{2}
+  \bibitem{paxos}
+    Paxos!
+
+\end{thebibliography}
+
 \end{document}

### 2012-05-04 paper/paper.tex
Fixing bib
Bluh bloh blah
More paper
No
Bloof
libmotmot intro and more bibses

@@
 \documentclass{sig-alternate}
+
+\usepackage{xspace}
+\newcommand\Motmot{\textsc{Motmot}\xspace}
 
@@
 
-\title{Motmot: A Secure, Distributed Chat Service}
+\title{Motmot: A Distributed Chat Protocol}
 
@@
   %
-\and
   \alignauthor
@@
   %
+  \and
   \alignauthor
@@
 
-We should write an abstract here.
+\Motmot is a new distributed text chat protocol based on Paxos. Unlike previous
+chat protocols like XMPP, OSCAR, and IRC, \Motmot provides guaranteed message
+delivery, total message ordering, end-to-end encryption, peer authentication,
+and message deniability.
+% TODO: write more
 
@@
 
-Wooooo.
+\section{libmotmot}
+
+Guaranteed delivery and total message ordering are perhaps the most fundamental
+chat features provided by \Motmot.  In order to obtain these properties, the
+clients participating in a chat must reach a consensus concerning the reception
+and order of chat events.  \Motmot assumes a benign failure model and uses a
+variant of the Paxos consensus protocol \cite{paxos}, which is the core of its
+client library \texttt{libmotmot}, in order to provide these guarantees.
+
+\subsection{Design}
+
+\subsection{Implementation}
+
+\subsection{Testing}
+
+\subsection{Lessons Learned}
+
+\section{Discovery Server}
+
+\section{Pidgin Plugin}
 
@@
 
-\begin{thebibliography}{2}
+\thebibliography{2}{
+  \bibitem{paxlive}
+    T. Chandra, R. Griesemer, and J. Redstone.
+    Paxos made live---an engineering perspective.
+    In \emph{Proc. 26th Symp. on Principles of Distributed Computing}, pp. 398-407, Aug. 2007.
+
   \bibitem{paxos}
-    Paxos!
+    L. Lamport.
+    The part-time parliament.
+    \emph{ACM Transactions on Computer Systems}, 16(2):133-169, 1998.
 
-\end{thebibliography}
+  \bibitem{paxsimp}
+    L. Lamport.
+    Paxos made simple.
+    \emph{ACM SIGACT News 32}, 4:18-25, Dec. 2001.
+
+  \bibitem{rfc}
+    P. Resnick, ed.
+    Internet Message Format, RFC 2822.
+    April 2001.
+}
 

### 2012-05-04 paper/paper.tex
bit more of the paper done

@@
 
+To allow our protocol to be utilized via Pidgin, we needed our plugin to interface with two major aspects of the protocol: the discovery server, and the libmotmot library.
+
+\subsection{Plugin Interactions with Discovery Server}
+
+\subsection{Plugin Interactions with Libmotmot}
+There were a few major difficulties that had to be addressed when calling libmotmot within the plugin:
+
+\begin{itemize}
+\item libpurple makes a distinction between "IMs" (two-person chats) and "chats" (three or more people); libmotmot does not.
+libpurple takes in different arguments for its IM-related and its chat-related functions
+
+\item One cannot modify the main Pidgin event loop from within a plugin (without patching libpurple itself).
+Instead, callbacks are made to plugin functions from within the main event loop, so finding a way to call libmotmot functions correctly (i.e.
+still honoring the correct control flow from libmotmot's perspective) and consistently within the bounds of these plugin functions was a second challenge.
+
+\end{itemize}
+
 \section{Conclusions}

### 2012-05-04 plugin/motmot.c
libmotmot doesn't work, but discovery server works

@@ -128,7 +128,24 @@ typedef struct {
   gpointer userdata;
 } GcFuncData;
 
+gint check_info(gconstpointer a, gconstpointer data){
+  int id = *((int *) data);
+  const MotmotInfo *info = a;
+  if(id == info -> id){
+    return 0;
+  }
+  return 1;
+}
 
+// finds an info
+MotmotInfo *find_info(motmot_conn *conn, int id){
+  GList *res = g_list_find_custom(conn -> info_list, &id, check_info);
+  if(res == NULL){
+    return NULL;
+  }
+
+  return res -> data;
+}
 // callbacks to momotinit
 // enter function
 void *enter(void *data){
@@ -137,6 +154,7 @@ void *enter(void *data){
   info -> internal_data = data;
 
   PurpleConnection *gc = GLOBAL_ACCOUNT -> gc;
+  motmot_conn *proto = gc -> proto_data;
   while(purple_find_chat(gc, chat_id) != NULL){
     chat_id += 1;
   }
@@ -147,7 +165,16 @@ void *enter(void *data){
 
   info -> id = id;
 
+  proto -> info_list = g_list_prepend(proto -> info_list, info);
+
   return info;
+}
+
+// leaving functions
+void leave_cb(void *data){
+  MotmotInfo *info = data;
+
+  serv_chat_leave(GLOBAL_ACCOUNT -> gc, info -> id);
 }
 
 /*
@@ -170,17 +197,18 @@ typedef struct {
 } GOfflineMessage;
 
 // Called by print_part_motmot to indicate successful chat departure.
+/*
 static void left_chat_room(PurpleConvChat *from, PurpleConvChat *to,
                            int id, const char *room, gpointer userdata) {
   if (from != to) {
-    /*  tell their chat window that we left */
     purple_debug_info("purplemot", "%s sees that %s left chat room %s\n",
                       to->nick, from->nick, room);
     purple_conv_chat_remove_user(to,
                                  from->nick,
-                                 NULL);  /* user-provided message, IRC style */
+                                 NULL);
   }
 }
+*/
 /**
  * query_status - query the status of a buddy
  *
@@ -352,6 +380,7 @@ static PurpleConnection *get_purplemot_gc(const char *username) {
 }
 
 // Called by print_chat_motmot, after paxos has said message is OK to send
+/*
[…]

### 2012-05-04 paper/paper.tex
Discovery server

@@
 and message deniability.
+% TODO: server federates, etc.
+% TODO: no xml! It's a feature
 % TODO: write more
@@
 \section{Discovery Server}
+
+The discovery server performs several tasks in the system: authentication and
+identity federation; presence management; and connection brokering.
+
+\subsection{Identity}
+
+Much like other federated protocols like XMPP and email, identity handles are
+based upon DNS. Every domain name with a SRV record can act as a \Motmot
+identity realm, and give out email-like addresses (an \verb`addr-spec` as
+defined in RFC 2822) for \Motmot communication.
+
+To grant a user an identity on a realm, the user first presents a username and
+password to the server, after which the server delegates authority for that name
+to the client by signing that client's key. This key-signing process is the
+method by which peers can authenticate securely to each other without a
+centralized broker.
+
+\subsection{Presence}
+
+In chat protocols, presence is the system by which users can specify a list of
+buddies to publish their online status to. The \Motmot discovery server provides
+a presence lookup service, by recording the status of all of the clients
+connected to it. Servers then push presence updates to all of the other servers
+on any user's buddy list, which in turn push the notification to the relevant
+clients.
+
+This subsystem system does not need to be robust or reliable, since all pushed
+state can be recreated from other sources, and since asynchronously pushed
+presence notifications do not need to be perfectly reliable to ensure a usable
+user interface.
+
+\subsection{Brokering}
+
+Finally, the server must perform message brokering. For clients behind firewalls
+or routers which perform network address translation (NAT), direct connections
+are not possible. In such a case, the discovery server might be forced to assist
+in connecting the clients.
+
+Technologies like STUN, TURN, and ICE provide implementations of ``NAT poking,''
+and while mechanisms for interfacing with such servers were considered in the
+design of the server, performing brokering itself was outside the scope of this
+version of the protocol.
+
+\subsection{Implementation}
+
+The \Motmot discovery server was implemented in Python using several open source
+libraries. The server itself is a custom RPC-like reactor/dispatcher written
+% TODO: cite gevent
+using gevent, a coroutine-based networking library. For our communication
+% TODO: cite msgpack
+protocol, we use msgpack to serialize messages. This allows for easy
+cross-language communication with clients and other servers, since
+implementations are available for most popular languages.
+
+We found that these two libraries provided good abstractions for the task of
+implementing a server. Gevent's coroutines provided an extremely useful
+concurrency model: they were executed in a single thread, so synchronization
+around global objects was not necessary, while avoiding the complexity (and
+unnecessary indentation) of a callback-based system. Msgpack also provided an
+excellent abstraction, allowing the server to assume that arbitrary data
+structures could be written to and read from the wire. The ad-hoc nature of the
+format also helped speed development of the server and protocol.
+
+% TODO: cite pyopenssl
+Finally, we used openssl (and the pyOpenSSL library in particular) to perform
+certificate operations
+
+\subsection{Lessons Learned}
+
+% TODO: EJ?
+>>>>>>> Discovery server
 

### 2012-05-04 server/install_test.sh
small modification to the testing install script

@@ -5,3 +5,4 @@
 sqlite3 config.db "INSERT INTO users(userName, password) values ('ej@bensing.com','12345');"
 
 sqlite3 config.db "INSERT INTO users(userName, password) values ('ej@bensing2.com','12345');"
+sqlite3 config.db "INSERT INTO users(userName, password) values ('ej@bensing1.com','12345');"

### 2012-05-04 paper/paper.tex
Oops #mergefail
.
More Paxos design

@@
 
-\usepackage{xspace}
+\usepackage{url,xspace}
 \newcommand\Motmot{\textsc{Motmot}\xspace}
+\newcommand\libmotmot{\texttt{libmotmot}\xspace}
 
@@
 and order of chat events.  \Motmot assumes a benign failure model and uses a
-variant of the Paxos consensus protocol \cite{paxos}, which is the core of its
-client library \texttt{libmotmot}, in order to provide these guarantees.
+variant of the Paxos consensus protocol \cite{paxos,paxsimp}, which is the core
+of its client library \libmotmot, in order to provide these guarantees.
 
 \subsection{Design}
+
+\libmotmot implements a slightly simplified variant of the Multi-Paxos
+algorithm, augmented by a number of out-of-band yet nonetheless indispensible
+side protocols.  This core Paxos implementation is wrapped by a client
+interface and uses MessagePack \cite{msgpack} for network communication.
+As with the other components of \Motmot, \libmotmot makes use of the GNOME
+Library (GLib) \cite{glib}.
+
+The Paxos implementation of \libmotmot makes a fundamental simplification to
+the standard algorithm: we do not allow participants who leave the protocol, by
+failure or otherwise, to rejoin.  In the original Paxos protocol, protocol
+participants may fail benignly or depart the protocol at any time, and return
+at some later point to continue voting and learning values in the system.
+This is necessary in order for participants to maintain a notion of persistent
+identity.  It is also needed to ensure recovery in the event of total system
+failure.  Participants keep a persistent log of the values committed in the
+protocol.  In the event of total failure, they can individually come back up,
+rejoin the protocol, and corroborate on their individual copies of the log,
+thereby allowing the protocol to continue where it left off.
+
+A chat protocol, however, has no need for this mode of recovery, nor for this
+particular notion of identity.  The persistent identity which users want from
+a chat service can be maintained out-of-band.  If a user leaves a chat, there
+is no need, at the level of the consensus protocol, to treat the user as the
+\emph{same} user if they returns to the chat later.  Users need not be informed
+of the messages that were exchanged in their absence.\footnote{We have not yet
+investigated whether or not a distributed chat protocol is able to support
+offline messages cleanly.}  So long as the system is able to determine, in a
+consistent manner, that the user has left, the user can gain a new identity
+within the protocol upon his or her return.
+
+Similarly, if a chat session ends, either due to failure or simply because all
+users leave, the termination is allowed to be final.  The participants in the
+chat may start a new session if they wish, but this new session need not have
+any memory nor knowledge of the old one.
+
+This simplification to Paxos yields many benefits.  For instance, clients may
+still choose to keep a record of chats in a local file, but by disallowing
+protocol rejoins, this persistent storage does not play any role in the safety
+or progress of \libmotmot's Paxos implementation.  We can also avoid the need
+for complicated and high-bandwidth recovery mechanisms.
+
+\libmotmot also implements a number of out-of-band protocols which support
+the core Paxos algorithm.  We will now describe each of these side protocols.
 
@@
 libraries. The server itself is a custom RPC-like reactor/dispatcher written
-% TODO: cite gevent
-using gevent, a coroutine-based networking library. For our communication
-% TODO: cite msgpack
-protocol, we use msgpack to serialize messages. This allows for easy
-cross-language communication with clients and other servers, since
-implementations are available for most popular languages.
+using GEvent \cite{glib}, a coroutine-based networking library. For our
+communication protocol, we use MessagePack \cite{msgpack} to serialize messages.
+This allows for easy cross-language communication with clients and other servers,
+since implementations are available for most popular languages.
 
@@
 Finally, we used openssl (and the pyOpenSSL library in particular) to perform
-certificate operations
+certificate operations.
 
@@
 % TODO: EJ?
->>>>>>> Discovery server
 
@@
 
-\item One cannot modify the main Pidgin event loop from within a plugin (without patching libpurple itself).
-Instead, callbacks are made to plugin functions from within the main event loop, so finding a way to call libmotmot functions correctly (i.e.
-still honoring the correct control flow from libmotmot's perspective) and consistently within the bounds of these plugin functions was a second challenge.
+\item The main Pidgin event loop cannot be modified from within a plugin (without patching libpurple itself).
+Instead, callbacks are made to plugin functions from within the main event loop, so we had to find a way to call libmotmot functions correctly (i.e.
+still honoring the correct control flow from libmotmot's perspective) and consistently within the bounds of these plugin functions
+
+\item libmotmot and libpurple use different handles to differentiate between chats and users, and translating between the two handles proved challenging, especially given the complexity of the information associated with libpurple chats (e.g.
+a single chat may have an id, a room name, PurpleConnection and PurpleConvChat objects...)
 
 \end{itemize}
+
+The bulk of the integration work was writing callbacks for the motmot\_init() function of libmotmot.
+Once these callbacks are written correctly, libmotmot handles most of the work of starting, sending, and coordinating chats automatically.
+These callbacks were, in almost all instances, responsible for calling libpurple functions, which handle the UI aspect of the associated libmotmot calls.
+In order to pass relevant information to our callbacks, we created a MotmotInfo struct, a large data struct that is used differently in different callbacks to get all the necessary info to the libpurple calls.
+
+We faced a challenge when writing the enter() function.
+This function is a callback responsible for allowing a user to enter a chat.
+Due to the structure of libmotmot, the only information passed into enter() as an argument is an internal handle, and we need create some sort of external handle.
+We solved this by creating a global variable, GLOBAL\_ACCOUNT, which represents the account that a user first logs in under, and we associate our external handle with this GLOBAL\_ACCOUNT.
+Though this ``works'' if a user only signs in with on eaccount, this breaks when a user decides to sign in with multiple accounts at once (something that libpurple allows), and should be addressed in future versions.
 
@@
 
+  \bibitem{glib}
+    GNOME Library.
+    \url{http://developer.gnome.org/about/}, 2011.
+
   \bibitem{paxos}
@@
 
+  \bibitem{msgpack}
+    MessagePack.
+    \url{http://www.msgpack.org/}, 2010.
+
   \bibitem{rfc}

### 2012-05-04 paper/paper.tex
Reflow Julie's stuff

--- 2012-05-04 paper/paper.tex
@@
 \newcommand\libmotmot{\texttt{libmotmot}\xspace}
+\newcommand\libpurple{\texttt{libpurple}\xspace}
 
@@
 
-To allow our protocol to be utilized via Pidgin, we needed our plugin to interface with two major aspects of the protocol: the discovery server, and the libmotmot library.
+To allow our protocol to be utilized via Pidgin, we needed our plugin to
+interface with two major aspects of the protocol: the discovery server, and the
+\libmotmot library.
 
@@
 
-\subsection{Plugin Interactions with Libmotmot}
-There were a few major difficulties that had to be addressed when calling libmotmot within the plugin:
+% TODO: Jeff?
+
+\subsection{Plugin Interactions with \libmotmot}
+There were a few major difficulties that had to be addressed when calling
+\libmotmot within the plugin:
 
 \begin{itemize}
-\item libpurple makes a distinction between "IMs" (two-person chats) and "chats" (three or more people); libmotmot does not.
-libpurple takes in different arguments for its IM-related and its chat-related functions
+\item \libpurple makes a distinction between "IMs" (two-person chats) and
+"chats" (three or more people); \libmotmot does not.  \libpurple takes in
+different arguments for its IM-related and its chat-related functions
 
-\item The main Pidgin event loop cannot be modified from within a plugin (without patching libpurple itself).
-Instead, callbacks are made to plugin functions from within the main event loop, so we had to find a way to call libmotmot functions correctly (i.e.
-still honoring the correct control flow from libmotmot's perspective) and consistently within the bounds of these plugin functions
+\item The main Pidgin event loop cannot be modified from within a plugin
+(without patching \libpurple itself).  Instead, callbacks are made to plugin
+functions from within the main event loop, so we had to find a way to call
+\libmotmot functions correctly (i.e. still honoring the correct control flow
+from \libmotmot's perspective) and consistently within the bounds of these
+plugin functions
 
-\item libmotmot and libpurple use different handles to differentiate between chats and users, and translating between the two handles proved challenging, especially given the complexity of the information associated with libpurple chats (e.g.
-a single chat may have an id, a room name, PurpleConnection and PurpleConvChat objects...)
+\item \libmotmot and \libpurple use different handles to differentiate between
+chats and users, and translating between the two handles proved challenging,
+especially given the complexity of the information associated with \libpurple
+chats (e.g. a single chat may have an id, a room name, \verb`PurpleConnection`
+and \verb`PurpleConvChat` objects...)
 
@@
 
-The bulk of the integration work was writing callbacks for the motmot\_init() function of libmotmot.
-Once these callbacks are written correctly, libmotmot handles most of the work of starting, sending, and coordinating chats automatically.
-These callbacks were, in almost all instances, responsible for calling libpurple functions, which handle the UI aspect of the associated libmotmot calls.
-In order to pass relevant information to our callbacks, we created a MotmotInfo struct, a large data struct that is used differently in different callbacks to get all the necessary info to the libpurple calls.
+The bulk of the integration work was writing callbacks for the
+\verb`motmot_init` function of \libmotmot.  Once these callbacks are written
+correctly, \libmotmot handles most of the work of starting, sending, and
+coordinating chats automatically.  These callbacks were, in almost all
+instances, responsible for calling \libpurple functions, which handle the UI
+aspect of the associated \libmotmot calls.  In order to pass relevant
+information to our callbacks, we created a \verb`MotmotInfo` struct, a large
+data struct that is used differently in different callbacks to get all the
+necessary info to the \libpurple calls.
 
-We faced a challenge when writing the enter() function.
-This function is a callback responsible for allowing a user to enter a chat.
-Due to the structure of libmotmot, the only information passed into enter() as an argument is an internal handle, and we need create some sort of external handle.
-We solved this by creating a global variable, GLOBAL\_ACCOUNT, which represents the account that a user first logs in under, and we associate our external handle with this GLOBAL\_ACCOUNT.
-Though this ``works'' if a user only signs in with on eaccount, this breaks when a user decides to sign in with multiple accounts at once (something that libpurple allows), and should be addressed in future versions.
+We faced a challenge when writing the \verb`enter` function.  This function is a
+callback responsible for allowing a user to enter a chat.  Due to the structure
+of \libmotmot, the only information passed into \verb`enter` as an argument is
+an internal handle, and we need create some sort of external handle.  We solved
+this by creating a global variable, \verb`GLOBAL_ACCOUNT`, which represents the
+account that a user first logs in under, and we associate our external handle
+with this \verb`GLOBAL_ACCOUNT`.  Though this ``works'' if a user only signs in
+with on eaccount, this breaks when a user decides to sign in with multiple
+accounts at once (something that \libpurple allows), and should be addressed in
+future versions.
 

### 2012-05-04 paper/paper.tex
Fix things

@@
 \libmotmot implements a slightly simplified variant of the Multi-Paxos
-algorithm, augmented by a number of out-of-band yet nonetheless indispensible
+algorithm, augmented by a number of out-of-band yet nonetheless indispensable
 side protocols.  This core Paxos implementation is wrapped by a client
@@
 around global objects was not necessary, while avoiding the complexity (and
-unnecessary indentation) of a callback-based system. Msgpack also provided an
+unnecessary indentation) of a callback-based system. MessagePack also provided an
 excellent abstraction, allowing the server to assume that arbitrary data
@@
 % TODO: cite pyopenssl
-Finally, we used openssl (and the pyOpenSSL library in particular) to perform
-certificate operations.
+Finally, we used \texttt{openssl} (and the \texttt{pyOpenSSL} library in
+particular) to perform certificate operations.
 

### 2012-05-04 paper/paper.tex
updated the paper

@@
 \libmotmot library.
+A libpurple protocol plugin is compiled as a statically linked library and installed into the lib directory
+of Pidgin. Upon start-up Pidgin loads each of the protocol plugins in its protocol directory. If the protocol is successfully loaded, the user
+can then create accounts of that specific protocol.
+In order to implement a protocol, the programmer must implement a certain number of functions specifying the code to run when the user takes
+an action, for example logging-in or joining a chat, in addition to listening
+for input over a socket. Implemented functions are stored in a struct. Calls to libmotmot and message passing between the client and the discovery server had to be
+implemented within these functions.
 
 \subsection{Plugin Interactions with Discovery Server}
+The procedure for hooking up the plugin to the discovery server is similar to the procedures in other chat protocols. Two user-info fields were
+added to the account creation menu, giving the server and port of the discovery server. In the login code, a connection with the discovery server
+is initiated, and once the connection is initiated an input listener is added and authentication credentials are sent. Once authentication succeeds,
+status updates and friend requests are handled. We used message pack in order to have a standard message format across multiple platforms..
+
+The input listener listens for input from the discovery server. Upon receiving a message it deserializes the message from the message pack buffer, finds
+the opcode, and calls the appropriate libpurple functions based on the opcode and the rest of the data sent by the message. Writes to the discovery server
+are handled in individual purple functions. Based on the user action, a message pack object is created, and the data is sent to the server across
+the connection. Messages are asynchronous.
 

### 2012-05-04 plugin/motmot.c

@@ -734,14 +734,14 @@ static void purplemot_login(PurpleAccount *acct)
   char **userparts;
   motmot_conn *conn;
   int port;
-
+/* This doesn't really work.
   if(GLOBAL_ACCOUNT != NULL){
 		purple_connection_error_reason (acct -> gc,
 			PURPLE_CONNECTION_ERROR_INVALID_SETTINGS,
 			_("You can't login to more than one motmot account"));
 		return;
   }
-
+*/
   GLOBAL_ACCOUNT = acct;
 
 	const char *username = purple_account_get_username(acct);
@@ -756,7 +756,7 @@ static void purplemot_login(PurpleAccount *acct)
   // initialize paxos functions
   // right now this does nothing, since these are do-nothing functions:
   // TODO fill out later
-  motmot_init(connect_motmot, print_chat_motmot, print_join_motmot, print_part_motmot, enter, leave_cb);
+  // motmot_init(connect_motmot, print_chat_motmot, print_join_motmot, print_part_motmot, enter, leave_cb);
 
 	if (strpbrk(username, " \t\v\r\n") != NULL) {
 		purple_connection_error_reason (gc,
@@ -930,7 +930,9 @@ static void motmot_parse(char *buffer, int len, PurpleConnection *gc){
   motmot_buddy *proto;
 
   msgpack_object_array ar = deser_get_array(buffer, len, &error);
+  purple_debug_info("motmot", "parsing data");
   if(error == PURPLEMOT_ERROR){
+    purple_debug_info("motmot", "error");
     return;
   }
   if(ar.size <= 0){
@@ -938,6 +940,7 @@ static void motmot_parse(char *buffer, int len, PurpleConnection *gc){
   }
 
   opcode = deser_get_pos_int(ar, 0, &error);
+  purple_debug_info("motmot", "opcode %d", opcode);
   if(error == PURPLEMOT_ERROR){
     return;
   }
@@ -1049,6 +1052,7 @@ static void motmot_parse(char *buffer, int len, PurpleConnection *gc){
       update_remote_status(a, friend_name, status);
 
       if(opcode == PUSH_FRIEND_ACCEPT){
+        purple_debug_info("motmot", "querying the status");
         query_status(friend_name, conn);
         break;
       }
@@ -1635,7 +1639,7 @@ static void purplemot_chat_leave(PurpleConnection *gc, int id) {
     return;
   }
   // disconnect
-  motmot_disconnect(info -> internal_data);
[…]

### 2012-05-04 server/install.sh

@@ -2,8 +2,8 @@
 
 #This script bootstraps a new sqlite db for testing
 sqlite3 config.db "CREATE TABLE users(userId INTEGER PRIMARY KEY AUTOINCREMENT, userName TEXT, password TEXT);"
-#sqlite3 config.db "INSERT INTO users(userName,password) values('jeff@motmottest.com','12345');"
-#sqlite3 config.db "INSERT INTO users(userName,password) values('jeff2@motmottest.com','12345');"
+sqlite3 config.db "INSERT INTO users(userName,password) values('jeff@motmottest.com','12345');"
+sqlite3 config.db "INSERT INTO users(userName,password) values('jeff2@motmottest.com','12345');"
 sqlite3 config.db "CREATE TABLE friends(friendId INTEGER PRIMARY KEY AUTOINCREMENT, userName TEXT, friend TEXT, accepted BOOLEAN);"
 mkdir cert
 mkdir ctest

### 2012-05-04 server/rpc.py

@@ -62,6 +62,7 @@ def writeback(conn, fn, args):
         print "Exception Occured on connection from {0}:{1}".format(conn.address[0], conn.address[1])
         rVal = [RM.BAD_STATUS,"Not a Valid Status Code"]
     finally:
+        print rVal
         conn.send(rVal)
 
 def rpc_dispatcher(conn):

### 2012-05-04 paper/paper.tex
More design

@@
 
-\subsection{Design}
+\subsection{Simplification}
 
@@
 or progress of \libmotmot's Paxos implementation.  We can also avoid the need
-for complicated and high-bandwidth recovery mechanisms.
+for complicated and high-bandwidth recovery mechanisms.  In \libmotmot's Paxos,
+as in the final protocol presented in the original Paxos paper, only one
+participant is designated as the proposer, and hence allowed to issue prepares
+and subsequently decrees, at any given time.  In the absence of rejoins, we
+are able to provide a very useful system invariant which full Paxos cannot
+have:
+\begin{quote}
+  If a proposer successfully completes a prepare phase, that proposer will
+  remain the protocol's designated proposer until it leaves the protocol, by
+  failure or otherwise.
+\end{quote}
+We discuss this invariant further in the following subsection.
 
-\libmotmot also implements a number of out-of-band protocols which support
-the core Paxos algorithm.  We will now describe each of these side protocols.
+\subsection{Support Protocols}
+
+In order to make Paxos efficient and usable, the core protocol needed to be
+supplemented with additional protocols which, despite being out-of-band, were
+no less integral to the functionality of \Motmot.  We will describe each of
+these side protocols in detail.
+
+\begin{description}
+  \item[Proposer election:]
+
+  \item[Greet protocol:]
+
+  \item[Out-of-band requests:]
+
+  \item[Reconnection:]
+
+  \item[Commit retry:]
+
+  \item[Log synchronization and truncation:]
+
+  \item[Connection continuations:]
+\end{description}
 
@@
 \subsection{Lessons Learned}
+
+Our simplification to Paxos is a lesson in the importance of end-to-end design
+\cite{end2end}.  The \libmotmot library is a backend protocol layer, on top of
+which chat semantics are implemented.  Without taking the semantics of the
+system's endpoints---the chat clients---into consideration, our consensus
+algorithm would have been unnecessarily featureful and complicated.
+
+The importance of out-of-band effort in distributed protocols was also made
+very clear to us in our implementation.  Other implementors of Paxos have also
+discovered the substantial amount of support needed to make the the protocol
+work in practical production-level settings \cite{paxlive}.  Clean designs
+and interfaces for these additional protocols was crucial in keeping our
+system simple, functional, and understandable.
 
@@
 
-\thebibliography{2}{
+\thebibliography{7}{
   \bibitem{paxlive}
@@
     April 2001.
+
+  \bibitem{end2end}
+    J. H. Saltzer, D. P. Reed, and D. D. Clark.
+    End-to-end arguments in system design.
+    \emph{ACM Transactions on Computer Systems}, 2(4):277-288, Nov. 1984.
 }

### 2012-05-04 paper/paper.tex
clarify chat/im

@@
 \begin{itemize}
-\item \libpurple makes a distinction between "IMs" (two-person chats) and
-"chats" (three or more people); \libmotmot does not.  \libpurple takes in
+\item \libpurple makes a distinction between ``IMs'' and
+``chats''; \libmotmot does not.
+(In particular, IMs are defined as a direct two-person communication only, whereas chats are an overarching category that can involve one, two, or more users.) \libpurple takes in
 different arguments for its IM-related and its chat-related functions

### 2012-05-04 server/rpc.py
added to libpurple section

@@ -62,7 +62,6 @@ def writeback(conn, fn, args):
         print "Exception Occured on connection from {0}:{1}".format(conn.address[0], conn.address[1])
         rVal = [RM.BAD_STATUS,"Not a Valid Status Code"]
     finally:
-        print rVal
         conn.send(rVal)
 
 def rpc_dispatcher(conn):

### 2012-05-04 plugin/motmot.c
updated Makefile for plugin

@@ -13,7 +13,9 @@
  * and friend requests. Actual chat is handled by the distributed paxos
  * algorithm.
  *
- * TODO describe how the paxos algorithm is hooked up
+ * Calls to the libmomot api are added in relevant user action functions.
+ * In addition event listeners are added to listen for relevant libmotmot
+ * actions.
  *
  * In addition to supporting presence and chatting, the protocol interacts
  * with the UI through calls to libpurple functions.
@@ -993,12 +995,10 @@ static void motmot_parse(char *buffer, int len, PurpleConnection *gc){
           update_remote_status(a, friend_name, status);
         }
         else{
-          // HACK
           proto = g_new0(motmot_buddy, 1);
           proto -> addr = addr;
           proto -> port = port;
           bud -> proto_data = proto;
-          // /HACK
           update_remote_status(a, friend_name, status);
           g_free(friend_name);
         }
@@ -1047,13 +1047,12 @@ static void motmot_parse(char *buffer, int len, PurpleConnection *gc){
       }
       status = deser_get_pos_int(ar, 2, &error);
       if(error == PURPLEMOT_ERROR){
+        g_free(friend_name);
         return;
       }
       update_remote_status(a, friend_name, status);
 
       if(opcode == PUSH_FRIEND_ACCEPT){
-        purple_debug_info("motmot", "querying the status");
-        query_status(friend_name, conn);
         break;
       }
 

### 2012-05-04 paper/paper.tex
added to libpurple section
Stuff

@@
 
-\usepackage{url,xspace}
+\usepackage{enumitem,url,xspace}
 \newcommand\Motmot{\textsc{Motmot}\xspace}
@@
 \newcommand\libpurple{\texttt{libpurple}\xspace}
+
+\setdescription{leftmargin=0pt}
 
@@
 \begin{description}
-  \item[Proposer election:]
+  \item[Join, part, and election:]
+    The original Paxos paper remarks on, but does not actually describe, a
+    procedure for electing a Paxos system's designated proposer; nor does it
+    describe mechanisms for adding and removing individuals from the system.
+    These are all, however, necessary to the functionality of our consensus
+    algorithm.  Election is necessary in order to ensure that the system makes
+    progress past the prepare phase of Paxos, lest multiple participants
+    attempt repeatedly to prepare over one another.  Join and part are required
+    in order to support any usable notion of chat.
+
+    In \libmotmot, we guarantee the consistency of joins, parts, and election
+    results by baking them in to the standard commit protocol.
 
@@
 data struct that is used differently in different callbacks to get all the
-necessary info to the \libpurple calls.
+necessary info to the \libpurple calls. \verb`MotmotInfo` contained the info
+necessary to mediate between \libpurple and \libmotmot. It contained both the
+id of the chat which was necessary for \libpurple and the opaque data handle
+necessary for \libmotmot.
 

### 2012-05-04 paper/paper.tex
Introduction

@@
 \section{Introduction}
+
+We wrote \Motmot due to our frustrations with existing chat protocols. XMPP, the
+protocol underlying Google Chat, is a federated XML-based chat protocol. Most
+annoyingly, we found that XMPP did not guarantee message delivery---when one
+of the users' connection dies, messages sent around the time of the
+disconnection may be lost, without any indication in the protocol that this has
+happened. Furthermore, XMPP messages are not totally ordered, and different chat
+participants can have different chat logs based on when messages were received.
+
+Finally, XMPP was an unsatisfying protocol technologically. While the protocol
+itself is federated, all chat messages are still sent to central servers for
+relaying. In addition to the obvious single-point-of-failure concerns, there are
+also several product-level concerns. In particular, any server which relays your
+messages is able to read the entire contents of the chat. Furthermore, your only
+guarantee that you are in fact talking to who you think you are comes from the
+servers you are talking to.
+
+Other chat protocols besides XMPP share similar properties. Skype, another
+popular chat protocol, acts in a peer-to-peer manner.\footnote{Recent rumors
+claim that Skype, under Microsoft's kindly direction, has moved some of the
+formerly distributed parts of the Skype protocol to a centrally managed server
+cluster. In light of such developments, it's unclear just how peer-to-peer
+Skype remains.} However, the protocol (and its primary implementation) are
+closed-source, so its guarantees cannot be further examined. As firm believers
+in sunlight and open protocols, we found Skype also wanting.
+
+\subsection{Design Goals}
+
+We set out implementing \Motmot with several design goals in mind.
+
+\begin{description}
+    \item[Multi-party] Chat sessions can occur between any number of
+    individuals.
+    \item[Reliability] All chat messages should be reliably delivered. Any message
+    recorded in the chat history is guaranteed to be delivered to the rest of the
+    \Motmot session.
+    \item[Total Ordering] All chat participants will receive the same messages
+    in exactly the same order.
+    \item[Security] All connections in the system will be encrypted with TLS,
+    using certificates (and client certificates) to authenticate the other party
+    to the communication.
+    \item[Deniability] While participants in the chat are confident that the
+    messages they receive are sent by the individuals specified, chat
+    participants cannot prove the provenance of any message to any third party.
+    In other words, after the chat has completed, all participants can fabricate
+    any equivalently valid chat history.
+    \item[Federation] Anyone can host their own server and issue authentication
+    credentials. Any \Motmot user can talk to any other \Motmot user (with an
+    identity in any realm) who correctly implements the protocol.
+    \item[Presence] Users can maintain buddy lists, and know the connection
+    status (online, busy, idle, etc.) of any buddy.
+\end{description}
 

### 2012-05-04 paper/paper.tex
Fix Jeff.

@@
 around global objects was not necessary, while avoiding the complexity (and
-unnecessary indentation) of a callback-based system. MessagePack also provided an
-excellent abstraction, allowing the server to assume that arbitrary data
+unnecessary indentation) of a callback-based system. MessagePack also provided
+an excellent abstraction, allowing the server to assume that arbitrary data
 structures could be written to and read from the wire. The ad-hoc nature of the
@@
 interface with two major aspects of the protocol: the discovery server, and the
-\libmotmot library.
-A libpurple protocol plugin is compiled as a statically linked library and installed into the lib directory
-of Pidgin. Upon start-up Pidgin loads each of the protocol plugins in its protocol directory. If the protocol is successfully loaded, the user
-can then create accounts of that specific protocol.
-In order to implement a protocol, the programmer must implement a certain number of functions specifying the code to run when the user takes
-an action, for example logging-in or joining a chat, in addition to listening
-for input over a socket. Implemented functions are stored in a struct. Calls to libmotmot and message passing between the client and the discovery server had to be
-implemented within these functions.
+\libmotmot library.  A \libpurple protocol plugin is compiled as a statically
+linked library and installed into the lib directory of Pidgin. Upon start-up
+Pidgin loads each of the protocol plugins in its protocol directory. If the
+protocol is successfully loaded, the user can then create accounts of that
+specific protocol.  In order to implement a protocol, the programmer must
+implement a certain number of functions specifying the code to run when the user
+takes an action, for example logging-in or joining a chat, in addition to
+listening for input over a socket. Implemented functions are stored in a struct.
+Calls to \libmotmot and message passing between the client and the discovery
+server had to be implemented within these functions.
 
 \subsection{Plugin Interactions with Discovery Server}
-The procedure for hooking up the plugin to the discovery server is similar to the procedures in other chat protocols. Two user-info fields were
-added to the account creation menu, giving the server and port of the discovery server. In the login code, a connection with the discovery server
-is initiated, and once the connection is initiated an input listener is added and authentication credentials are sent. Once authentication succeeds,
-status updates and friend requests are handled. We used message pack in order to have a standard message format across multiple platforms..
+The procedure for hooking up the plugin to the discovery server is similar to
+the procedures in other chat protocols. Two user-info fields were added to the
+account creation menu, giving the server and port of the discovery server. In
+the login code, a connection with the discovery server is initiated, and once
+the connection is initiated an input listener is added and authentication
+credentials are sent. Once authentication succeeds, status updates and friend
+requests are handled. We used message pack in order to have a standard message
+format across multiple platforms.
 
-The input listener listens for input from the discovery server. Upon receiving a message it deserializes the message from the message pack buffer, finds
-the opcode, and calls the appropriate libpurple functions based on the opcode and the rest of the data sent by the message. Writes to the discovery server
-are handled in individual purple functions. Based on the user action, a message pack object is created, and the data is sent to the server across
-the connection. Messages are asynchronous.
-
-% TODO: Jeff?
+The input listener listens for input from the discovery server. Upon receiving a
+message it deserializes the message from the message pack buffer, finds the
+opcode, and calls the appropriate \libpurple functions based on the opcode and
+the rest of the data sent by the message. Writes to the discovery server are
+handled in individual purple functions. Based on the user action, a message pack
+object is created, and the data is sent to the server across the connection.
+Messages are asynchronous.
 
@@
 \begin{itemize}
-\item \libpurple makes a distinction between ``IMs'' and
-``chats''; \libmotmot does not.
-(In particular, IMs are defined as a direct two-person communication only, whereas chats are an overarching category that can involve one, two, or more users.) \libpurple takes in
-different arguments for its IM-related and its chat-related functions
+\item \libpurple makes a distinction between ``IMs'' and ``chats''; \libmotmot
+does not.  (In particular, IMs are defined as a direct two-person communication
+only, whereas chats are an overarching category that can involve one, two, or
+more users.) \libpurple takes in different arguments for its IM-related and its
+chat-related functions
 

### 2012-05-04 plugin/motmot.c
updated plugin doc

@@ -397,6 +397,12 @@ static void receive_chat_message(PurpleConvChat *from, PurpleConvChat *to,
 */
 // returns fd on success, -1 on failure—>TODO(Julie) fix this once MAX does his shit
 // it'll also be a GIOChannel * then.  So yeah.
+
+/**
+ * connectSuccess - Callback for when a p2p connection is established for libmotmot.
+ *
+ * @param data - User data.
+ */
 static void connectSuccess(gpointer data, gint source, const gchar *error_message)
 {
   //MotmotInfo *info = data;
@@ -1364,7 +1370,13 @@ static void purplemot_change_passwd(PurpleConnection *gc, const char *old_pass,
 static gint motmot_node_cmp(const char *a, const char *b){
   return strcmp(a, b);
 }
-
+/**
+ * purple_mot_add_buddy - code called when a buddy is added in libpurple
+ *
+ * @param gc The account connection
+ * @param buddy The buddy to be added
+ * @param group The group (we don't use this)
+ */
 static void purplemot_add_buddy(PurpleConnection *gc, PurpleBuddy *buddy,
                                PurpleGroup *group)
 {
@@ -1380,6 +1392,9 @@ static void purplemot_add_buddy(PurpleConnection *gc, PurpleBuddy *buddy,
   msgpack_packer *pk = msgpack_packer_new(buffer, msgpack_sbuffer_write);
   GList *el = g_list_find_custom(conn -> acceptance_list, 
     purple_buddy_get_name(buddy), (GCompareFunc) motmot_node_cmp);
+
+  // This is a hack so that the acceptance to a friend request is only sent once the buddy is added on our end,
+  // not when the add is authorized.
   if(el != NULL){
     name = el -> data;
     msgpack_pack_array(pk, 2);
@@ -1398,7 +1413,8 @@ static void purplemot_add_buddy(PurpleConnection *gc, PurpleBuddy *buddy,
     
     return;
   }
-  
+
+  // send out the request
   msgpack_pack_array(pk, 2);
   msgpack_pack_int(pk, REGISTER_FRIEND);
   msgpack_pack_raw(pk, strlen(purple_buddy_get_name(buddy)));

### 2012-05-04 paper/paper.tex
Fix an hbox
Random shit

@@
 \begin{itemize}
-\item \libpurple makes a distinction between ``IMs'' and ``chats''; \libmotmot
+\item \libpurple makes a distinction between ``chats'' and ``IMs''; \libmotmot
 does not.  (In particular, IMs are defined as a direct two-person communication
@@
 with this \verb`GLOBAL_ACCOUNT`.  Though this ``works'' if a user only signs in
-with on eaccount, this breaks when a user decides to sign in with multiple
+with one account, this breaks when a user decides to sign in with multiple
 accounts at once (something that \libpurple allows), and should be addressed in

### 2012-05-04 paper/paper.tex
Conclusion

@@
 
-Wheeee.
+\Motmot is a novel distributed chat protocol that offers features not present in
+other chat protocols. Its reliance on the Paxos algorithm allow it to react
+properly to the changing topology of chatting peers, and is the core improvement
+on other protocols.
+
+While we have not had a chance to test the protocol in practice, we believe that
+\Motmot not only provides interesting technical guarantees, but is also usable
+in practice.
+
+In summary, we believe that \Motmot is the best thing since sliced bread.
 

### 2012-05-04 paper/paper.tex
edited in ej's changes

@@
 
+We rewrote the server several times because we did not understand a few of the technical requirements of our design choices; namely, that communication needs to be asynchronous.
+Asynchrony allows for  push notifications to go over the same socket as all other requests.
+Once we understood that communication needed to by asynchronous, the next  design choice was how to implement socket reading.
+In an earlier implementation, we were inadvertently throwing away parts of messages, causing incorrect behavior.
+Moving to the unpacker object in the msgpack library and using its feed() function solved this problem and prevented data loss.
+
+From there, the implementation was fairly straight forward.
+Servers currently authenticate by checking that the connection is coming from the IP that DNS records have for the domain.
+This could be made more robust (and is perhaps a concern for future versions), but as of the present, it allows for servers to be easily added to the global network and communicate.
+
+Multiplexing over a single socket was by far the most difficult aspect of this server.
+Many of the problems with the original synchronous implementation could have been solved by adding a second socket, but we decided that using a single socket is superior because it has simpler failure modes.
+
+We quickly learned that writing good test code for distributed systems is both difficult and time consuming.
+The server test code, while not exhaustive, covers many common situations.
+However, testing for partial failure of messages and other network related failures will be implemented in a future version.
+Additionally, in a future version, we plan to add message IDs to each call so that the client can know which response corresponds to which request.
+
 % TODO: EJ?

### 2012-05-04 paper/paper.tex
updated status report in paper

@@
 object is created, and the data is sent to the server across the connection.
-Messages are asynchronous.
+Messages are asynchronous. Status updates also include the address and port
+of a user so the client can know the location of a user to connect to.
+
+The code to interact with the discovery server currently has a bug involving
+status updates after accepting a friend request.
 
@@
 future versions.
+
+We were unable to fully integrate the \libmotmot protocol in the plugin.
+The code is outlined with in the plugin file, however it is not currently
+functional. Hopefully this will be fixed in future versions so that
+\libmotmot can be used as a \libpurple plugin.
 

### 2012-05-04 plugin/motmot.c

@@ -401,7 +401,9 @@ static void receive_chat_message(PurpleConvChat *from, PurpleConvChat *to,
 /**
  * connectSuccess - Callback for when a p2p connection is established for libmotmot.
  *
- * @param data - User data.
+ * @param data  User data.
+ * @param source The file descriptor for the socket, -1 on error.
+ * @param error_message Error message
  */
 static void connectSuccess(gpointer data, gint source, const gchar *error_message)
 {
@@ -415,7 +417,13 @@ static void connectSuccess(gpointer data, gint source, const gchar *error_messag
   return;
 }
 
-
+/**
+ * connect_motmot - Connect callback for motmot_init to initiate a function
+ *
+ * @param desc username
+ * @param len length of the username
+ * @param motmot_connect_cb continuation necessary for libmotmot
+ */
 int connect_motmot(const void *desc, size_t len, struct motmot_connect_cb *cb)
 {
   PurpleBuddy *bud = purple_find_buddy(GLOBAL_ACCOUNT, (const char *) desc);
@@ -433,6 +441,15 @@ int connect_motmot(const void *desc, size_t len, struct motmot_connect_cb *cb)
 //                        const char *message, PurpleMessageFlags flags)
 
 // libmotmot callback. Calls receive_chat_message.
+/**
+ * print_chat_motmot - callback for receiving a chat, prints message
+ *
+ * @param message the message
+ * @param len length of the message
+ * @param desc user who sent it
+ * @param size length of username's message
+ * @param data MotmotInfo struct to identify chat to libpurple
+ */
 int print_chat_motmot(const void *message, size_t len, void *desc, size_t size, void *data)
 {
   MotmotInfo *minfo = data;

### 2012-05-04 server/install.sh

@@ -2,8 +2,8 @@
 
 #This script bootstraps a new sqlite db for testing
 sqlite3 config.db "CREATE TABLE users(userId INTEGER PRIMARY KEY AUTOINCREMENT, userName TEXT, password TEXT);"
-sqlite3 config.db "INSERT INTO users(userName,password) values('jeff@motmottest.com','12345');"
-sqlite3 config.db "INSERT INTO users(userName,password) values('jeff2@motmottest.com','12345');"
+#sqlite3 config.db "INSERT INTO users(userName,password) values('jeff@motmottest.com','12345');"
+#sqlite3 config.db "INSERT INTO users(userName,password) values('jeff2@motmottest.com','12345');"
 sqlite3 config.db "CREATE TABLE friends(friendId INTEGER PRIMARY KEY AUTOINCREMENT, userName TEXT, friend TEXT, accepted BOOLEAN);"
 mkdir cert
 mkdir ctest

### 2012-05-04 paper/paper.tex
changed a word
libmotmot testing

@@
 
+Testing \libmotmot in a realistic setting was also challenging. We resorted to
+client ``fuzzing'' as our primary testing strategy, in which we randomly issued
+valid commands to a pool of connected \Motmot instances. These commands included
+messages, join invitations, and parts, and were issued at random intervals to
+random chat participants.
+
+One of the biggest challenges was writing a testing framework that allowed us to
+script interactions between many processes. Our first attempt was a Ruby DSL
+that allowed small scripts to spawn and interact with a given number of \Motmot
+instances. After performing arbitrary manipulations to the client pool,
+including sending messages, creating new clients, and killing clients, a reaper
+thread ensures message ordering.
+
+The second iteration of a testing framework instead relied more heavily on UNIX
+abstractions. First, we create a UNIX named pipe to act as \Motmot's input
+channel (replacing \texttt{stdin}). A small utility program
+(\texttt{test/motmot}) then assists in \texttt{dup2}-ing the correct file
+descriptors in place and spawning the correct process. This allows many
+\Motmot's to be spawned in separate windows, while centralizing the control of
+the instances. Unlike the previous testing framework, \texttt{stdout} is not
+checked. Finally, a central trace runner writes commands to the appropriate
+named pipes, causing messages to be sent to various windows.
+
+A key debugging feature of both of these frameworks was the ability to attach
+debuggers to every \Motmot instance spawned. This served to be a critical
+debugging tool, especially when analyzing segmentation faults or ensuring client
+state consistency.
+
+One final problem when testing \libmotmot was in simulating the network.  Chats
+occur over TCP in the real world, which have some amount of latency and can
+perform packet fragmentation, neither of which is a concern when interfacing
+over UNIX domain sockets. To simulate a lossy network (on which connections
+occasionally die), we wrote a network proxy, which ferried data from one UNIX
+socket to another, fragmenting them and adding between 5 and 500 milliseconds of
+latency. Connections were also occasionally dropped at some random interval.
+
+While simulating a network proved difficult, we found that fuzzing with network
+delays was a very effective way of finding bugs. After sending tens of millions
+of messages through \libmotmot, we are fairly confident that it correctly
+implements Paxos and delivers chats reliably, even at very high message
+throughputs (on the order of a thousand messages a second on low-latency
+networks).
+
 \subsection{Lessons Learned}
@@
 functional. Hopefully this will be fixed in future versions so that
-\libmotmot can be used as a \libpurple plugin.
+\libmotmot can be used within a \libpurple plugin.
 

### 2012-05-04 plugin/motmot.c
added more plugin data

@@ -20,6 +20,9 @@
  * In addition to supporting presence and chatting, the protocol interacts
  * with the UI through calls to libpurple functions.
  *
+ * In general functions that are documented with a javadoc style are
+ * functions we implemented. The rest are mostly functions from
+ * nullprpl.
  */
 
 #include <stdarg.h>
@@ -934,7 +937,13 @@ static void deny_cb(void *data){
   
 }
 
-
+/**
+ * motmot_parse - The main function for handling data sent from the discovery server.
+ *
+ * @param buffer The data from the discovery server.
+ * @param len The length of the buffer
+ * @param gc The connection
+ */
 
 static void motmot_parse(char *buffer, int len, PurpleConnection *gc){
   char *friend_name;
@@ -1166,11 +1175,18 @@ static void motmot_input_cb(gpointer *data, PurpleSslConnection *gsc, PurpleInpu
 		return;
 	}
 
-  motmot_parse(buffer, l, gc);
+  motmot_parse(buffer, read_bytes, gc);
   g_free(buffer);
   return;
 }
 
+/**
+ * motmot_login_cb - callback for successful connection
+ *
+ * @param data The connection
+ * @param gsc The ssl connection
+ * @param cond The input condition
+ */
 
 static void motmot_login_cb(gpointer data, PurpleSslConnection *gsc, PurpleInputCondition cond){
   purple_debug_info("motmot", "connection achieved");
@@ -1187,6 +1203,13 @@ static void motmot_login_cb(gpointer data, PurpleSslConnection *gsc, PurpleInput
 	}
 }
 
+/**
+ * motmot_login_failure - callback for a failed connection
+ *
+ * @param gsc The ssl connection
+ * @param error The error type
+ * @param data User supplied data (the connection)
+ */
 
 static void
 motmot_login_failure(PurpleSslConnection *gsc, PurpleSslErrorType error,
@@ -1200,7 +1223,11 @@ motmot_login_failure(PurpleSslConnection *gsc, PurpleSslErrorType error,
 	purple_connection_ssl_error (gc, error);
 }
 
-
+/**
+ * purplemot_close - called when libpurple shuts down an account
+ *
+ * @params gc The connection
+ */
 
 
 static void purplemot_close(PurpleConnection *gc)
@@ -1331,6 +1358,13 @@ static void purplemot_get_info(PurpleConnection *gc, const char *username) {
                          NULL);     /* userdata for callback */
 }
 
+/**
[…]

### 2012-05-04 paper/paper.tex
Most of design

@@
 or progress of \libmotmot's Paxos implementation.  We can also avoid the need
-for complicated and high-bandwidth recovery mechanisms.  In \libmotmot's Paxos,
+for complicated and high-volume recovery mechanisms.  In \libmotmot's Paxos,
 as in the final protocol presented in the original Paxos paper, only one
@@
   \item[Join, part, and election:]
-    The original Paxos paper remarks on, but does not actually describe, a
+    The original Paxos paper mentions, but does not actually provide, a
     procedure for electing a Paxos system's designated proposer; nor does it
@@
     In \libmotmot, we guarantee the consistency of joins, parts, and election
-    results by baking them in to the standard commit protocol.
+    results by baking them in to the standard commit protocol.  Each protocol
+    participant keeps a list of all the participants in the chat, ordered
+    by session-unique identifiers.  Participants can be added and removed from
+    this list only through Paxos commits decreeing either \verb`DEC_JOIN` or
+    \verb`DEC_PART`, which ensures the consistency of the participants in this
+    list and their order.  Participants are identified by the ID of the Multi-%
+    Paxos instance which decreed their join.
+
+    Proposer election is performed locally, with each participant simply
+    recognizing the earliest-joined participant to whom they have a live socket
+    connection as the designated proposer.  Participants will only accept a
+    prepare phase issued by this designated proposer.  When a new proposer
+    successfully completes its prepare phase, it broadcasts a decree requesting
+    that all participants evict the old proposer from their participant lists.
+    In this manner, we are able to guarantee the invariant described above.
+
+    This mechanism may occasionally result in the forced eviction of a live
+    participant who was the previous proposer.  It is conceivable that transient
+    network problems might cause a majority of participants to lose their
+    connection to the original proposer, and hence to accept the prepare of
+    the new proposer.  Before the new proposer is able to issue its decree
+    parting the old proposer, however, these connections may all be
+    reestablished, and the old proposer may be booted even though the system
+    as a whole currently recognizes its presence.  This causes no loss of
+    consistency and merely forces an old proposer in this situation to rejoin
+    the chat.  We consider this a worthwhile tradeoff given the usefulness
+    of our invariant.
 
   \item[Greet protocol:]
+    The presence of joins and parts demands a mechanism for initiating
+    communication with new protocol participants.  When a join is successfully
+    accepted by a majority, the proposer will decree it.  Subsequently, the
+    proposer will send the new participant a welcome message, which transfers
+    all of its protocol state to the new participant.  The proposer always has
+    the most up-to-date state, and since the network guarantees FIFO message
+    ordering and the proposer is the only one who can issue commands to modify
+    protocol state, the new participant's state will be consistent.  The
+    proposer also assigns the new participant its ID (i.e., the instance
+    number of its join decree).
+
+    The new participant will then open sockets to each participant on the
+    list it receives from the proposer.  It will then send each participant
+    a hello message.  This hello message is also used in the reconnect protocol
+    and is used to bind sockets to the protocol state objects representing
+    participants.
 
   \item[Out-of-band requests:]
+    The proposer is the only protocol participant able to order and to decree
+    a chat message.  Hence, all participants must have a mechanism for asking
+    the proposer to decree a message.  These requests also apply for joins
+    (though not to parts, since we do not support forced kicking of users from
+    a chat).
+
+    We want to avoid stressing the network, and in particular want to keep the
+    volume of data the proposer must handle to a minimum, since the proposer is
+    one of the system's primary bottlenecks.  Thus, we did not want to force
+    the proposer to broadcast the arbitrarily-large chat messages produced by
+    the various users in the chat.  Instead of sending along the actual message,
+    the requester of a message generates a session-unique identifier for the
+    message, consisting of the requester's own participant ID and a locally
+    unique request ID.  The requester then takes on the responsibility of
+    broadcasting this identifier along with the actual message contents to all
+    participants.  When this message is received, each participant caches it
+    locally.  The proposer will make a decree which contains the identifier of
+    the request; when the other participants receive a commit for the decree,
+    they can do a local lookup for the message contents.
+
+    This has the added benefit of providing all participants an assurance that
+    message contents come directly from their sources without the possibility
+    of corruption by the proposer.  At worst, the proposer can modify the
+    cache index of the message.
+
+    Because our request mechanism is done out-of-band, commits for a given
+    request may arrive after the message contents are received from the initial
+    broadcast.  We implement a basic protocol for retrieving and resending
+    these messages if they cannot be found.
 
@@
 
-  \item[Commit retry:]
+  \item[In-order learn:]
+    \libmotmot's clients are notified of Paxos commits via a set of callbacks.
+    In order to ensure the total ordering of chats as seen by the user, we
+    ``learn'' about commits in order of Paxos instance number.  This means that
+    a missing commit message could prevent client-side progress.  To rectify
+    this, we implement a straightforward means for any participant to ask the
+    proposer to remind them of a committed value.
 
   \item[Log synchronization and truncation:]
+    Commits made in Paxos must be kept by all protocol participants in order
+    to obtain Paxos's correctness guarantees.  Since we keep all of our logs
+    in memory, however, we need a mechanism for pruning information which
+    the entire system has finished learning.
+
+    We implement an out-of-band sync-and-truncate protocol, which is used by
+    the proposer to issue consistent system-wide deletion of old commits.  A
+    proposer asks all protocol participants to return the last contiguous
+    commit which their respective clients have learned about.  Once it receives
+    a response from \emph{all} participants, it can take the minimum instance
+    number of those commits and safely command all participants to truncate
+    up to that point.
 

### 2012-05-05 paper/paper.tex
Finishing paper

@@
 
-  \item[Reconnection:]
+  \item[Redirection:]
+    In the presence of connection failures, protocol participants may at times
+    have an incorrect notion of who the designated proposer is.  In particular,
+    we have the following three cases:
+    \begin{enumerate}
+      \item A participant thinks it is the proposer and tries to prepare, but
+        a majority of participants are still connected to the true proposer.
+      \item A proposer thinks that a participant has been disconnected from
+        the system, but a majority of participants are still connected to it.
+      \item A participant issues a request to another participant who it
+        incorrectly identifies as the proposer.
+    \end{enumerate}
+    To deal with the former two cases, we implement a majority rejection
+    mechanism for prepares and part decrees.  When a majority rejects a
+    prepare, they indicate their identification of the true proposer.
+    The preparer, who belives himself to be the proposer, then attempts
+    to reconnect to this true proposer.  Similarly, rejections of part
+    induce the issuer of the decree to attempt reconnection to the target
+    of the part.
+
+    If participants misidentify the proposer when making a request, the
+    individual that they identified as the proposer will return the true
+    identity of the proposer.  A single such response is enough to cause
+    the requester to attempt reconnection.
 
@@
     up to that point.
-
-  \item[Connection continuations:]
 \end{description}
-
-\subsection{Implementation}
 
@@
 
-\section{Acknowledgments}
-
 \thebibliography{7}{

### 2012-11-12 6.858/proposal/proposal.tex
Stub out project proposal

\documentclass[12pt]{article}

\usepackage{geometry, xspace}
\usepackage[parfill]{parskip}
\geometry{letterpaper}

\newcommand{\trill}{\textsc{Trill}\xspace}

\title{Trill: Secure firewall-traversing networking}
\author{Carl Jackson and Max Wang}
\date{November 5, 2012}

\begin{document}

\maketitle

\section*{Overview}
We propose to design and implement \trill, a networking abstraction that
provides secure communication between federated identities, either of which
could be behind a firewall.

\section*{Security}
% TODO

\section*{Federated Identity}
% TODO

\section*{Firewall Traversal}
% TODO

\section*{Networking Abstractions}
% TODO

\end{document}

### 2012-11-12 6.858/proposal/proposal.tex
Security!

@@
 \section*{Security}
-% TODO
+The security of our system will be based on standard cryptographic
+infrastructure and libraries. By ``identities,'' we mean public key pairs, the
+private half of which is uniquely known by an individual. This identity can be
+given one or more names by means of X.509 certificates, and standard mechanisms
+like TLS allow owners of names to securely establish a communication channel and
+mutually verify the other's identity. Due to implementation restrictions in our
+target environment, we will be using DTLS, a younger protocol that uses UDP as
+its transport layer, and we will be using the implementation of DTLS provided by
+the open source GnuTLS library.
 

### 2012-11-12 6.858/proposal/proposal.tex
Formatted some things

@@
 
-\usepackage{geometry, xspace}
+\usepackage{fancyhdr, geometry, xspace}
 \usepackage[parfill]{parskip}
-\geometry{letterpaper}
+\geometry{letterpaper,text={7in,9in},centering}
+\pagestyle{fancy}
 
-\newcommand{\trill}{\textsc{Trill}\xspace}
+\newcommand{\Trill}{\textsc{Trill}\xspace}
 
-\title{Trill: Secure firewall-traversing networking}
+\title{\Trill: Secure firewall-traversing networking}
 \author{Carl Jackson and Max Wang}
 \date{November 5, 2012}
+
+\makeatletter
+\renewcommand{\maketitle}{%
+  \thispagestyle{plain}%
+  \begin{center}%
+    {\LARGE \@title \par}%
+    {\large \@author \par}%
+    {\large \@date \par}%
+  \end{center}%
+}
+\makeatother
 
@@
 \section*{Overview}
-We propose to design and implement \trill, a networking abstraction that
+We propose to design and implement \Trill, a networking abstraction that
 provides secure communication between federated identities, either of which
@@
 given one or more names by means of X.509 certificates, and standard mechanisms
-like TLS allow owners of names to securely establish a communication channel and
-mutually verify the other's identity. Due to implementation restrictions in our
-target environment, we will be using DTLS, a younger protocol that uses UDP as
-its transport layer, and we will be using the implementation of DTLS provided by
-the open source GnuTLS library.
+like TLS allow owners of names to securely establish a communication channel
+and mutually verify the other's identity. Due to implementation restrictions in
+our target environment, we will be using DTLS, a younger protocol that uses UDP
+as its transport layer, and we will be using the implementation of DTLS
+provided by the open source GnuTLS library.
 

### 2012-11-12 6.858/proposal/proposal.tex
Delicious Italian food

@@
 \section*{Federated Identity}
-% TODO
+As with other federated protocols like XMPP and email, \Trill identities will
+be based on DNS.  Every domain name with a SRV record can act as a \Trill
+identity realm, supporting email address-like identities (an \verb!addr-spec!
+as defined in RFC 2822).
+
+To grant a user an identity on a realm, the user first presents a username and
+password to the server, after which the server delegates authority for that
+name to the client by signing that client's key. This key-signing process is
+the method by which peers can authenticate securely to each other without a
+centralized broker.
 

### 2012-11-12 6.858/proposal/proposal.tex
Firewalls!

@@
 \section*{Firewall Traversal}
-% TODO
+Due to restrictions on many firewalls and NATs in place on the Internet, it is
+impossible for arbitrary clients on the internet to establish peer-to-peer
+communication. There are techniques for firewall/NAT traversal which involve
+``tricking'' the networking layer into establishing a connection. We plan on
+reimplementing something similar to \texttt{chownat} (http://samy.pl/chownat/),
+which appears to be a standard way to do this.
 

### 2012-11-12 6.858/proposal/proposal.tex
A nice white sauce for the pasta

@@
 
-To grant a user an identity on a realm, the user first presents a username and
-password to the server, after which the server delegates authority for that
-name to the client by signing that client's key. This key-signing process is
-the method by which peers can authenticate securely to each other without a
-centralized broker.
+Each \Trill domain should host a federation server, which is able to delegate
+authority for identities to various clients.  Any client may present the server
+with an identity and a password for that identity.  If the client's
+authentication succeeds, the server signs the client's key and returns the
+resultant certificate.  Using these certificates, clients can then authenticate
+securely to one another without further use of the server.
 

### 2012-11-12 6.858/proposal/proposal.tex
Header woo woo woo
Networking!

@@
 \pagestyle{fancy}
+
+\fancyhead[RO]{Carl Jackson and Max Wang}
 
@@
 ``tricking'' the networking layer into establishing a connection. We plan on
-reimplementing something similar to \texttt{chownat} (http://samy.pl/chownat/),
-which appears to be a standard way to do this.
+reimplementing something similar to \texttt{chownat} (http://samy.pl/chownat/)
+for \Trill, which appears to be a standard way to do this.
 
 \section*{Networking Abstractions}
-% TODO
+While using UDP allows us to traverse firewalls, it does not present a
+particularly nice interface to the programmer. As always, those who do not
+understand TCP are doomed to reimplement it, poorly. We hope to sidestep this
+problem entirely by providing an abstraction more similar to SCTP: datagrams may
+be of arbitrary length, and will arrive correctly ordered in their entirety or
+not at all. However, datagrams may be received in any order relative to other
+datagrams. We believe that this is the right level of abstraction to expose for
+most peer-to-peer applications.
+
+If time permits, we will additionally provide TCP-like congestion control in
+\Trill.
 

### 2012-11-12 6.858/proposal/proposal.tex
URLifying

@@
 
-\usepackage{fancyhdr, geometry, xspace}
+\usepackage{fancyhdr, geometry, hyperref, xspace}
 \usepackage[parfill]{parskip}
+
 \geometry{letterpaper,text={7in,9in},centering}
@@
 \fancyhead[RO]{Carl Jackson and Max Wang}
+\fancyhead[LO]{\url{http://github.com/motmotchat/trill}}
 
@@
 ``tricking'' the networking layer into establishing a connection. We plan on
-reimplementing something similar to \texttt{chownat} (http://samy.pl/chownat/)
-for \Trill, which appears to be a standard way to do this.
+reimplementing something similar to \texttt{chownat}
+(\url{http://samy.pl/chownat/}) for \Trill, which appears to be a standard way
+to do this.
 

### 2012-12-14 plume/client.rb
Bluh, just exit on client EOF without cleanup

@@ -27,10 +27,7 @@ def prompt(conn, msg=nil)
 
   EM.defer do
     unless buf = Readline.readline('> ', true)
-      EM.next_tick do
-        conn.close_connection
-        abort "\n"
-      end
+      abort "\n"
     end
 
     input = buf.strip.split
@@ -47,10 +44,7 @@ def prompt(conn, msg=nil)
     when 'help', 'h', '?'
       msg = USAGE
     when 'exit', 'quit', 'q'
-      EM.next_tick do
-        conn.close_connection
-        exit
-      end
+      exit
     else
       msg = "Invalid command.  Type 'h' for help."
     end

### 2012-12-14 plume/plume.rb
Fixing PlumeServer args

@@ -59,7 +59,8 @@ class PlumeServer < PlumeConn
     return close_connection if addr.nil? or port.nil?
 
     # Route the connection request to the peer's Plume server.
-    EM.connect(addr, port, PlumeServer, key_file, crt_file, @conns) do |conn|
+    EM.connect(addr, port, PlumeServer,
+               key_file, crt_file, @conns, @udp_reqs) do |conn|
       conn.send_data ['route', [cert, peer, op, payload]].to_msgpack
     end
   end

### 2012-12-14 plume/conn/client_plume.rb
Moar little typos

@@ -26,7 +26,7 @@ class ClientPlumeConn < PlumeConn
   def initialize(key_file, crt_file)
     super key_file, crt_file
     @peers = {}
-    @upds = {}
+    @udps = {}
     @timers = {}
   end
 
@@ -51,8 +51,8 @@ class ClientPlumeConn < PlumeConn
 
     # If we don't know our UDP IP and port for this peer, ask the server.
     if not @udps[peer]
-      cookie = 97392   # TODO: make this random
-      @upds[cookie] = peer
+      cookie = 49897392   # TODO: make this random
+      @udps[cookie] = peer
 
       send_data ['udp', [cert.to_pem, cookie]].to_msgpack
       return

### 2012-12-14 plume/plume.rb

@@ -68,7 +68,7 @@ class PlumeServer < PlumeConn
   #
   # Register a UDP self-identification request.
   #
-  def udp(cookie)
+  def udp(cert, cookie)
     @udp_reqs[cookie] = self
     send_data ['ack_udp', [cookie]].to_msgpack
   end

### 2012-12-14 plume/conn/client_plume.rb
Whoops, making ops legal

@@ -20,7 +20,7 @@ end
 
 class ClientPlumeConn < PlumeConn
 
-  LEGAL_OPS = %w(connect identify cert)
+  LEGAL_OPS = %w(connect identify cert ack_udp udp)
   OP_PREFIX = 'recv_'
 
   def initialize(key_file, crt_file)

### 2012-12-14 plume/plume.rb

@@ -14,7 +14,7 @@ require_relative 'util.rb'
 
 class PlumeServer < PlumeConn
 
-  LEGAL_OPS = %w(route)
+  LEGAL_OPS = %w(route udp)
 
   def initialize(key_file, crt_file, conn_table, udp_reqs)
     super key_file, crt_file

### 2012-12-14 plume/conn/client_plume.rb
So bad at typing

@@ -109,7 +109,7 @@ class ClientPlumeConn < PlumeConn
 
     @timers[cookie] = EM::PeriodicTimer.new(0.5) do
       domain = parse_email(cert_cn(cert)).domain
-      addr, port = dns_get_srv("_plume-upd._udp.#{domain}")
+      addr, port = dns_get_srv("_plume-udp._udp.#{domain}")
 
       send_datagram cookie.to_msgpack, addr, port
     end

### 2012-12-14 plume/conn/plume_udp.rb
Lolol, !UDP.include? SSL

@@ -1,0 +1,30 @@
+#
+# plume_udp.rb - Plume UDP echo service listener.
+#
+
+require 'eventmachine'
+require 'msgpack'
+
+class PlumeUDPEcho < EM::Connection
+
+  def initialize(udp_reqs)
+    @udp_reqs = udp_reqs
+  end
+
+  def receive_data(buf)
+    begin
+      cookie = MessagePack.unpack(buf)
+    rescue MessagePack::UnpackError
+    end
+
+    return unless @udp_reqs[cookie]
+
+    # Send the client's IP and port back via TCP, then clear the cookie.
+    @udp_reqs[cookie].send_data ['udp', [cookie, get_peeraddr]].to_msgpack
+    @udp_reqs.delete cookie
+  end
+
+  def get_peeraddr
+    Socket.unpack_sockaddr_in(get_peername).reverse
+  end
+end

### 2012-12-14 plume/plume.rb

@@ -10,6 +10,7 @@ require 'msgpack'
 require 'openssl'
 
 require_relative 'conn/conn.rb'
+require_relative 'conn/plume_udp.rb'
 require_relative 'util.rb'
 
 class PlumeServer < PlumeConn
@@ -71,29 +72,6 @@ class PlumeServer < PlumeConn
   def udp(cert, cookie)
     @udp_reqs[cookie] = self
     send_data ['ack_udp', [cookie]].to_msgpack
-  end
-end
-
-#
-# Plume UDP echo service listener.
-#
-class PlumeUDPEcho < PlumeConn
-
-  def initialize(udp_reqs)
-    @udp_reqs = udp_reqs
-  end
-
-  def receive_data(buf)
-    begin
-      cookie = MessagePack.unpack(buf)
-    rescue MessagePack::UnpackError
-    end
-
-    return unless @udp_reqs[cookie]
-
-    # Send the client's IP and port back via TCP, then clear the cookie.
-    @udp_reqs[cookie].send_data ['udp', [cookie, get_peeraddr]].to_msgpack
-    @udp_reqs.delete cookie
   end
 end
 

### 2012-12-14 plume/conn/client_plume.rb
Finishing UDP echo service

@@ -2,6 +2,7 @@
 # client_plume.rb - Connection between a client and a Plume server.
 #
 
+require 'eventmachine'
 require 'msgpack'
 require 'openssl'
 
@@ -16,6 +17,10 @@ class PlumeID
   def initialize(cert)
     @cert = cert
   end
+end
+
+class UDPConn < EM::Connection
+  attr_accessor :ip, :port
 end
 
 class ClientPlumeConn < PlumeConn
@@ -59,7 +64,8 @@ class ClientPlumeConn < PlumeConn
     end
 
     # Sign our IP and port, encrypt the signature, and send it to our peer.
-    sig = OpenSSL::PKCS7.sign(cert, key, get_sockaddr.join(':'))
+    ip, port = @udps[peer].ip, @udps[peer].port
+    sig = OpenSSL::PKCS7.sign(cert, key, "#{ip}:#{port}")
     id_enc = @peers[peer].cert.public_key.public_encrypt_block(sig.to_pem)
 
     send_data ['route', [cert.to_pem, peer, 'connect', id_enc]].to_msgpack
@@ -107,11 +113,13 @@ class ClientPlumeConn < PlumeConn
   def recv_ack_udp(cookie)
     return unless @udps[cookie]
 
+    @udps[@udps[cookie]] = EM::open_datagram_socket '0.0.0.0', 0, UDPConn
+
     @timers[cookie] = EM::PeriodicTimer.new(0.5) do
       domain = parse_email(cert_cn(cert)).domain
       addr, port = dns_get_srv("_plume-udp._udp.#{domain}")
 
-      send_datagram cookie.to_msgpack, addr, port
+      @udps[@udps[cookie]].send_datagram cookie.to_msgpack, addr, port
     end
   end
 
@@ -119,8 +127,9 @@ class ClientPlumeConn < PlumeConn
   # Receive a response to a UDP self-identification request.
   #
   def recv_udp(cookie, ip_port)
-    # Match the peer with our UDP IP and port.
-    @udps[@udps[cookie]] = ip_port
+    # Match the UDP connection with our IP and port.
+    udp_conn = @udps[@udps[cookie]]
+    udp_conn.ip, udp_conn.port = ip_port
 
     # Clear relevant state.
     @timers[cookie].cancel

### 2012-12-14 plume/conn/client_plume.rb
Randomness

@@ -32,6 +32,8 @@ class ClientPlumeConn < PlumeConn
     super key_file, crt_file
     @peers = {}
     @udps = {}
+
+    @prng = Random.new
     @timers = {}
   end
 
@@ -56,7 +58,7 @@ class ClientPlumeConn < PlumeConn
 
     # If we don't know our UDP IP and port for this peer, ask the server.
     if not @udps[peer]
-      cookie = 49897392   # TODO: make this random
+      cookie = @prng.rand(1 << 64)
       @udps[cookie] = peer
 
       send_data ['udp', [cert.to_pem, cookie]].to_msgpack

### 2012-12-14 plume/conn/client_plume.rb
Use handle in place of cert_cn(cert)

@@ -118,7 +118,7 @@ class ClientPlumeConn < PlumeConn
     @udps[@udps[cookie]] = EM::open_datagram_socket '0.0.0.0', 0, UDPConn
 
     @timers[cookie] = EM::PeriodicTimer.new(0.5) do
-      domain = parse_email(cert_cn(cert)).domain
+      domain = parse_email(handle).domain
       addr, port = dns_get_srv("_plume-udp._udp.#{domain}")
 
       @udps[@udps[cookie]].send_datagram cookie.to_msgpack, addr, port

### 2012-12-14 src/gnutls.c
Free up resources better

@@ -62,6 +62,13 @@ trill_tls_free(struct trill_connection *conn)
 {
   assert(conn != NULL);
 
+  // Attempt to gracefully disconnect. Note that this could return an error code
+  // (GNUTLS_E_AGAIN, or maybe a real error), but we don't care about that,
+  // since we'll be dropping the underlying UDP transport in just a bit anyways.
+  gnutls_bye(conn->tc_tls.tt_session, GNUTLS_SHUT_RDWR);
+
+  gnutls_deinit(conn->tc_tls.tt_session);
+
   gnutls_certificate_free_credentials(conn->tc_tls.tt_creds);
 
   return 0;

### 2012-12-14 src/network.c

@@ -150,7 +150,11 @@ int
 trill_connection_free(struct trill_connection *conn)
 {
   int retval = 0;
+
   assert(conn != NULL && "Attempting to free a null connection");
+
+  retval = trill_tls_free(conn);
+
   if (conn->tc_sock_fd != -1) {
     if (close(conn->tc_sock_fd) == -1) {
       log_errno("Error closing socket");

### 2012-12-14 plume/conn/client_plume.rb
Finishing up connection protocol

@@ -80,13 +80,21 @@ class ClientPlumeConn < PlumeConn
   #
   def recv_connect(peer_cert, peer_id_enc)
     peer_cert = OpenSSL::X509::Certificate.new peer_cert
-    p7 = OpenSSL::PKCS7.new(key.private_decrypt_block(peer_id_enc))
+    peer = cert_cn(peer_cert)
 
+    p7 = OpenSSL::PKCS7.new(key.private_decrypt_block(peer_id_enc))
     ca_store = OpenSSL::X509::Store.new
     verify = p7.verify([peer_cert], ca_store, nil, OpenSSL::PKCS7::NOVERIFY)
+
     return close_connection unless verify
 
-    puts p7.data
+    # Connect to the peer if we haven't already.
+    # TODO: Maybe see if the user actually wants to connect.
+    if not @peers[peer]
+      @peers[peer] = PlumeID.new(peer_cert)
+      connect peer
+    end
+    @peers[peer].ip, @peers[peer].port = p7.data.split(':')
   end
 
   #

### 2012-12-14 src/common.h
I guess it might be nice to support NSS too
...but not now.

@@ -9,6 +9,8 @@
 #include "gnutls.h"
 #elif defined(TRILL_USE_OPENSSL)
 #error "OpenSSL not implemented yet!"
+#elif defined(TRILL_USE_NSS)
+#error "NSS not implemented yet!"
 #else
 #error "Unknown cryptographic library"
 #endif

### 2012-12-14 src/gnutls.c
Actually call the connected callback

@@ -197,6 +197,9 @@ trill_tls_handshake(struct trill_connection *conn)
   } else if (ret == 0) {
     log_info("TLS is established!");
     conn->tc_state = TC_STATE_ESTABLISHED;
+    if (conn->tc_connected_cb != NULL) {
+      conn->tc_connected_cb(conn->tc_event_loop_data);
+    }
   } else {
     log_error("Something went wrong with the TLS handshake");
   }

### 2012-12-14 src/network.c
More bad state errors

@@ -125,6 +125,11 @@ trill_connect(struct trill_connection *conn, const char *remote,
   assert(conn != NULL && "Attempting to connect with a null connection");
   assert(remote != NULL && "Connecting to a null address");
 
+  if (conn->tc_state != TC_STATE_INIT) {
+    log_error("Connection is in an inappropriate state!");
+    return -1;
+  }
+
   bzero(&addr, sizeof(addr));
   addr.sin_family = AF_INET;
   addr.sin_port = htons(port);

### 2012-12-14 src/gnutls.c
I hate endianness

@@ -258,6 +258,10 @@ trill_tls_can_read(struct trill_connection *conn)
       len = gnutls_record_recv_seq(conn->tc_tls.tt_session, buf, sizeof(buf),
           (unsigned char *)&seq);
 
+      // TODO: sequence number is in big-endian. It'd probably be more useful to
+      // everyone if it was converted into host endinanness, but be64toh is
+      // Linux only.
+
       assert(conn->tc_recv_cb != NULL && "No callback set");
       conn->tc_recv_cb(conn->tc_event_loop_data, buf, len, &seq);
       break;

### 2012-12-14 src/trill.h
Some documentation appears!

@@ -4,19 +4,86 @@
 #include <stdint.h>
 #include <sys/types.h>
 
+/**
+ * struct trill_connection - An opaque type representing a connection to another
+ * Trill client.
+ */
 struct trill_connection;
 
+/**
+ * trill_callback_t - Callback type for most Trill callbacks.
+ *
+ * @param conn      The connection upon which to act.
+ * @return          0 to cancel the callback, 1 to reschedule it.
+ */
 typedef int (*trill_callback_t)(struct trill_connection *conn);
-typedef int (*trill_want_write_callback_t)(void *data);
-typedef int (*trill_want_timeout_callback_t)(void *data,
+
+/**
+ * trill_want_write_callback_t - Callback from Trill to the application's event
+ * loop.
+ *
+ * Called when Trill wishes to receive can-write events.
+ *
+ * @param cb_data   The callback data pointer set with `trill_set_data`. In most
+ *                  cases, some represent of the event loop's representation of
+ *                  the connection object and other application-specific state.
+ * @return          0 on success, nonzero on error.
+ */
+typedef int (*trill_want_write_callback_t)(void *cb_data);
+
+/**
+ * trill_want_timeout_callback_t - Callback from Trill to the application's
+ * event loop.
+ *
+ * Called when Trill wishes to receive timeout events. These should fire at
+ * regular intervals until `callback` returns 0.
+ *
+ * @param cb_data   The callback data pointer set with `trill_set_data`. In most
+ *                  cases, some represent of the event loop's representation of
+ *                  the connection object and other application-specific state.
+ * @param callback  The Trill function to call periodically. The timer should be
+ *                  canceled when this callback returns 0.
+ * @param millis    The number of milliseconds between calls.
+ * @return          0 on success, nonzero on error.
+ */
+typedef int (*trill_want_timeout_callback_t)(void *cb_data,
     trill_callback_t callback, unsigned int millis);
 
+/**
+ * trill_connected_callback_t - Callback from Trill to the application's event
+ * loop.
+ *
+ * Called when Trill has established a secure connection and is ready to send
+ * and receive application data.
+ *
+ * @param cb_data   The callback data pointer set with `trill_set_data`. In most
+ *                  cases, some represent of the event loop's representation of
+ *                  the connection object and other application-specific state.
+ */
 typedef void (*trill_connected_callback_t)(void *cb_data);
-typedef ssize_t (*trill_recv_callback_t)(void *cb_data, void *data, size_t len, uint64_t *seq);
+
[…]

### 2012-12-14 src/plume.c
Uh, we can do DNS now in C-land

@@ -1,0 +1,35 @@
+/**
+ * plume.c - Plume client.
+ */
+
+#include <glib.h>
+#include <gio/gio.h>
+
+#include "log.h"
+
+GMainLoop *gmain;
+
+void
+plume_connect_server(GObject *resolver, GAsyncResult *res, void *data)
+{
+  GList *srvs;
+
+  srvs = g_resolver_lookup_service_finish((GResolver *)resolver, res, NULL);
+
+  log_info("%s:%hu", g_srv_target_get_hostname(srvs->data),
+      g_srv_target_get_port(srvs->data));
+}
+
+int
+main(int argc, char *argv[])
+{
+  g_type_init();
+
+  g_resolver_lookup_service_async(g_resolver_get_default(), "plume", "tcp",
+      "mxawng.com", NULL, plume_connect_server, NULL);
+
+  gmain = g_main_loop_new(g_main_context_default(), 0);
+  g_main_loop_run(gmain);
+
+  return 0;
+}

### 2012-12-14 src/common.h
Verify certs

@@ -47,6 +47,8 @@ struct trill_connection {
 
   uint32_t tc_server_priority[2];
 
+  char *tc_remote_user;
+
   trill_callback_t tc_can_read_cb;
   trill_callback_t tc_can_write_cb;
 

### 2012-12-14 src/gnutls.c

@@ -18,6 +18,7 @@ static gnutls_priority_t priority_cache;
 int trill_tls_can_read(struct trill_connection *conn);
 int trill_tls_can_write(struct trill_connection *conn);
 int trill_tls_handshake_retry(struct trill_connection *conn);
+int trill_tls_verify_cert(gnutls_session_t session);
 
 int
 trill_crypto_init(void)
@@ -53,6 +54,9 @@ trill_tls_init(struct trill_connection *conn)
     log_error("Error allocating crypto credentials");
     return 1;
   }
+
+  gnutls_certificate_set_verify_function(conn->tc_tls.tt_creds,
+      trill_tls_verify_cert);
 
   return 0;
 }
@@ -132,6 +136,8 @@ trill_start_tls(struct trill_connection *conn)
     return 1;
   }
 
+  gnutls_session_set_ptr(conn->tc_tls.tt_session, conn);
+
   if (gnutls_priority_set(conn->tc_tls.tt_session, priority_cache)) {
     log_error("Unable to set priorities on new session");
     return 1;
@@ -205,6 +211,17 @@ trill_tls_handshake(struct trill_connection *conn)
   }
 
   return 0;
+}
+
+int
+trill_tls_verify_cert(gnutls_session_t session)
+{
+  unsigned int errors;
+  struct trill_connection *conn = gnutls_session_get_ptr(session);
+
+  gnutls_certificate_verify_peers3(session, conn->tc_remote_user, &errors);
+
+  return errors == 0;
 }
 
 ssize_t

### 2012-12-14 src/main.c

@@ -116,7 +116,7 @@ main(int argc, char *argv[])
   }
   *buf_ptr++ = '\0';
 
-  trill_connect(conn, buf, atoi(buf_ptr));
+  trill_connect(conn, "testing.com", buf, atoi(buf_ptr));
   log_info("Connecting to %s on port %d", buf, atoi(buf_ptr));
 
   g_io_add_watch(myconn->chan, G_IO_IN, socket_can_read, conn);

### 2012-12-14 src/network.c

@@ -6,6 +6,7 @@
 #include <net/if.h>
 #include <stdint.h>
 #include <stdlib.h>
+#include <string.h>
 #include <strings.h>
 #include <sys/socket.h>
 #include <unistd.h>
@@ -116,8 +117,8 @@ trill_connection_new()
 }
 
 int
-trill_connect(struct trill_connection *conn, const char *remote,
-    uint16_t port)
+trill_connect(struct trill_connection *conn, const char *who,
+    const char *remote, uint16_t port)
 {
   int ret;
   struct sockaddr_in addr;
@@ -145,6 +146,7 @@ trill_connect(struct trill_connection *conn, const char *remote,
   connect(conn->tc_sock_fd, (struct sockaddr *) &addr, sizeof(addr));
 
   conn->tc_state = TC_STATE_PROBING;
+  conn->tc_remote_user = strdup(who);
 
   trill_want_timeout_callback(conn->tc_event_loop_data, trill_connection_probe, 1000);
 

### 2012-12-14 src/trill.h

@@ -123,13 +123,14 @@ int trill_connection_free(struct trill_connection *conn);
  * trill_connect - Connect to another Trill client.
  *
  * @param conn      The connection on which to connect.
+ * @param who       The remote hostname to connect to.
  * @param address   The remote address to connect to. Should be an IPv4 address
  *                  in presentation format (see `inet_pton`).
  * @param port      The remote port to connect to.
  * @return          0 on success, nonzero on error.
  */
-int trill_connect(struct trill_connection *conn, const char *address,
-    uint16_t port);
+int trill_connect(struct trill_connection *conn, const char *who,
+    const char *address, uint16_t port);
 
 /**
  * trill_get_fd - Get the file descriptor associated with a given connection.

### 2012-12-14 6.858/paper/paper.tex
Make the proposal past tense
This is a good first approximation of a final paper.

--- 2012-11-12 6.858/proposal/proposal.tex
@@
 
-\usepackage{fancyhdr, geometry, xspace}
+\usepackage{fancyhdr, geometry, hyperref, xspace}
 \usepackage[parfill]{parskip}
+
 \geometry{letterpaper,text={7in,9in},centering}
 \pagestyle{fancy}
+
+\fancyhead[RO]{Carl Jackson and Max Wang}
+\fancyhead[LO]{\url{http://github.com/motmotchat/trill}}
 
@@
 \author{Carl Jackson and Max Wang}
-\date{November 5, 2012}
+\date{December 14, 2012}
 
@@
 \section*{Overview}
-We propose to design and implement \Trill, a networking abstraction that
-provides secure communication between federated identities, either of which
-could be behind a firewall.
+We implemented \Trill, a networking abstraction that provides secure
+communication between federated identities, either of which could be behind a
+firewall.
 
 \section*{Security}
-The security of our system will be based on standard cryptographic
-infrastructure and libraries. By ``identities,'' we mean public key pairs, the
-private half of which is uniquely known by an individual. This identity can be
-given one or more names by means of X.509 certificates, and standard mechanisms
-like TLS allow owners of names to securely establish a communication channel
-and mutually verify the other's identity. Due to implementation restrictions in
-our target environment, we will be using DTLS, a younger protocol that uses UDP
-as its transport layer, and we will be using the implementation of DTLS
-provided by the open source GnuTLS library.
+The security of our system is based on standard cryptographic infrastructure and
+libraries. By ``identities,'' we mean public key pairs, the private half of
+which is uniquely known by an individual. This identity can be given one or more
+names by means of X.509 certificates, and standard mechanisms like TLS allow
+owners of names to securely establish a communication channel and mutually
+verify the other's identity. Due to implementation restrictions in our target
+environment, we used DTLS, a younger protocol that uses UDP as its transport
+layer; specifically, we used the implementation of DTLS provided by the open
+source GnuTLS library.
 
 \section*{Federated Identity}
-As with other federated protocols like XMPP and email, \Trill identities will
-be based on DNS.  Every domain name with a SRV record can act as a \Trill
-identity realm, supporting email address-like identities (an \verb!addr-spec!
-as defined in RFC 2822).
+As with other federated protocols like XMPP and email, \Trill identities are
+based on DNS.  Every domain name with a SRV record can act as a \Trill identity
+realm, supporting email address-like identities (an \verb!addr-spec!  as defined
+in RFC 2822).
 
-Each \Trill domain should host a federation server, which is able to delegate
+Each \Trill domain hosts a federation server, which is able to delegate
 authority for identities to various clients.  Any client may present the server
@@
 communication. There are techniques for firewall/NAT traversal which involve
-``tricking'' the networking layer into establishing a connection. We plan on
-reimplementing something similar to \texttt{chownat} (http://samy.pl/chownat/),
+``tricking'' the networking layer into establishing a connection. We implemented
+something similar to \texttt{chownat} (\url{http://samy.pl/chownat/}) in \Trill,
 which appears to be a standard way to do this.
 
-\section*{Networking Abstractions}
-% TODO
-
 \end{document}

### 2012-12-14 src/plume.c
More server connect and TLS support

@@ -9,15 +9,59 @@
 
 GMainLoop *gmain;
 
-void
-plume_connect_server(GObject *resolver, GAsyncResult *res, void *data)
+int
+plume_tls_verify_cert(GTlsConnection *conn, GTlsCertificate *cert,
+    GTlsCertificateFlags errors, void *data)
 {
-  GList *srvs;
+  log_warn("plume_tls_verify_cert");
 
-  srvs = g_resolver_lookup_service_finish((GResolver *)resolver, res, NULL);
+  return 1;
+}
 
-  log_info("%s:%hu", g_srv_target_get_hostname(srvs->data),
-      g_srv_target_get_port(srvs->data));
+void
+plume_tls_setup(GSocketClient *client, GSocketClientEvent event,
+    GSocketConnectable *addr, GIOStream *conn, void *data)
+{
+  GTlsCertificate *cert;
+  GTlsClientConnection *tls_conn;
+
+  log_warn("plume_tls_setup: %d", event);
+
+  if (event != G_SOCKET_CLIENT_TLS_HANDSHAKING) {
+    return;
+  }
+
+  tls_conn = (GTlsClientConnection *)conn;
+
+  cert = g_tls_certificate_new_from_file("client.crt", NULL);
+  g_tls_connection_set_certificate((GTlsConnection *)tls_conn, cert);
+  g_tls_client_connection_set_validation_flags(tls_conn, 0);
+
+  g_signal_connect(tls_conn, "accept-certificate",
+      (GCallback)plume_tls_verify_cert, NULL);
+}
+
+void
+plume_connect_server(GObject *obj, GAsyncResult *res, void *data)
+{
+  GResolver *resolver;
+  GSrvTarget *srv;
+
+  GSocketConnectable *addr;
+  GSocketClient *client;
+  GSocketConnection *conn;
+
+  resolver = (GResolver *)obj;
+  srv = g_resolver_lookup_service_finish(resolver, res, NULL)->data;
+
+  client = g_socket_client_new();
+  g_socket_client_set_tls(client, TRUE);
+  addr = g_network_address_new(g_srv_target_get_hostname(srv),
+      g_srv_target_get_port(srv));
+
+  g_signal_connect(client, "event", (GCallback)plume_tls_setup, NULL);
+
+  conn = g_socket_client_connect(client, addr, NULL, NULL);
 }
 
 int

### 2012-12-14 6.858/paper/paper.tex
Random thoughts on paper
Formatting changes

@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 
-\usepackage{fancyhdr, geometry, hyperref, xspace}
+\usepackage{fancyhdr, geometry, hyperref, sectsty, titlesec, xspace}
 \usepackage[parfill]{parskip}
@@
 
-\fancyhead[RO]{Carl Jackson and Max Wang}
+\sectionfont{\vspace{-2pt}\large}
+\subsectionfont{\vspace{-2pt}\normalsize}
+\subsubsectionfont{\normalfont\itshape}
+
+\titlespacing\section{0pt}{10pt}{2pt}
+\titlespacing\subsection{0pt}{8pt}{0pt}
+
+\fancyhead[RO]{Carl Jackson \& Max Wang}
 \fancyhead[LO]{\url{http://github.com/motmotchat/trill}}
@@
 \newcommand{\Trill}{\textsc{Trill}\xspace}
+\newcommand{\Plume}{\textsc{Plume}\xspace}
 
 \title{\Trill: Secure firewall-traversing networking}
-\author{Carl Jackson and Max Wang}
+\author{Carl Jackson \and Max Wang}
 \date{December 14, 2012}
-
-\makeatletter
-\renewcommand{\maketitle}{%
-  \thispagestyle{plain}%
-  \begin{center}%
-    {\LARGE \@title \par}%
-    {\large \@author \par}%
-    {\large \@date \par}%
-  \end{center}%
-}
-\makeatother
 
@@
 
-\section*{Overview}
 We implemented \Trill, a networking abstraction that provides secure
@@
 
-\section*{Security}
+\section{Design}
+
+\subsection{Security}
 The security of our system is based on standard cryptographic infrastructure and
@@
 names by means of X.509 certificates, and standard mechanisms like TLS allow
-owners of names to securely establish a communication channel and mutually
-verify the other's identity. Due to implementation restrictions in our target
-environment, we used DTLS, a younger protocol that uses UDP as its transport
-layer; specifically, we used the implementation of DTLS provided by the open
-source GnuTLS library.
+owners of identities to securely establish a communication channel and mutually
+verify the other's identity.
 
 \section*{Federated Identity}
-As with other federated protocols like XMPP and email, \Trill identities are
-based on DNS.  Every domain name with a SRV record can act as a \Trill identity
-realm, supporting email address-like identities (an \verb!addr-spec!  as defined
-in RFC 2822).
+As with other federated protocols like XMPP and email, \Trill handles are based
+on DNS. In particular, \Trill's handles use RFC 2822's \verb`addr-spec`
+specification, in which every \verb`domain` is allowed complete control over the
+\Trill handles given out as \verb`local-part`'s at their domain. Domains wishing
+to run \Trill must also publish information about its \Trill servers using DNS
+SRV records.
 
@@
 authentication succeeds, the server signs the client's key and returns the
-resultant certificate.  Using these certificates, clients can then authenticate
-securely to one another without further use of the server.
+resultant certificate.  Using these certificates, clients can then prove their
+handles securely to one another without further use of the server.
 
-\section*{Firewall Traversal}
+Note that the assigning of handles is orthogonal to the security of the protocol
+itself, and is only useful in associating human-friendly names to their actual
+cryptographic identity, i.e., their private key. Users who do not wish to use
+\Trill's federated identity system can instead self-sign their certificates and
+verify these credentials out-of-band. This is very similar to PGP's ``web of
+trust,'' and while we have not implemented such a system, we imagine the
+integration of \Trill with PGP would be trivial. An SSH-like Trust On First Use
+(TOFU) system might also be viable here.
+
+\section*{Establishing Trust}
+Trust in \Trill is established between two clients when each of them verifies
+that the other client's certificate is signed by the other client's domain key.
+However, the clients must also recursively verify this key, to make sure the
+purported domain key was actually issued by the controller of the domain name.
+
+This can be done in two ways. In the first, the user can use the traditional TLS
+mechanism of maintaining a set of ``root'' certificate authorities, which are
+securely preinstalled as part of the operating system or TLS library. The client
+can verify that the domain owner certificate has a chain of trust up to a root
+CA, which is sufficient to verify all end users at that domain. Note that the
+entire user-to-root certificate chain cannot be verified at once, since it is
+unlikely that the \Trill domain key is allowed to function as a certificate
+authority; instead, by convention, a certificate issued to a domain name is
+allowed by \Trill to subsequently sign the keys of users at that domain. One
+could imagine an extension to X.509 that properly codified this, and although
+writing RFCs is outside the scope of this project, we believe that such an
+extension would be useful for a wide variety of use cases (for instance, imagine
+an email service which was allowed to sign the keys of its users in this way).
+
+The second way trust can be established is through DNSSEC, which has very
+recently become mature enough to use in this way. In such a system, a domain's
+\Trill root key could be published as a DNSSEC-signed TXT record. A DNS client
+with the root zone signing keys can verify any DNSSEC-signed record recursively,
+and in this way, discover and authenticate any \Trill key it wishes.
+
+We believe this second mechanism is both more secure and more closely aligns
+with how we are using DNS in \Trill. Since root certificate authorities are
+allowed to sign keys for any common name, the compromise of a single certificate
+authority is sufficient to break the security of the entire system. In contrast,
+DNSSEC keys are hierarchical, such that the compromise of a single key
+(``\verb`.cn.`,'' say) cannot affect the security of unrelated parts of the system
+(``\verb`.com.`''), a very desirable property. Furthermore, since DNS (and
+DNSSEC) is the only way to establish ``ownership'' of a domain, it makes sense
+to fundamentally link trust in \Trill with trust in DNS. We consider any
+individual who is allowed to make DNSSEC-signed changes to a given domain's DNS
+records the ``owner'' of the domain, so a DNSSEC-rooted trust model is the most
+natural fit for \Trill.
+
+Unfortunately, while the \verb`.com.` realm is signed by DNSSEC, there are
+relatively few registrars that support it. While we hope the situation improves,
+it seems like CA-rooted trust is the most compatible system for implementing
+\Trill in the near future, and it is what we have used in our implementation.
+
+\subsection{Firewall Traversal}
 Due to restrictions on many firewalls and NATs in place on the Internet, it is
@@
 
+\section{Implementation}
+% trill - client library. allows plug-and-play event loops, architected to (one
+%   day!) support multiple TLS libraries.
+% plume - server. separate login and protocol servers. access to domain root
+%   keys has very small surface area. chroot more!
+
+\subsection{Trill client library}
+
+\subsection{Plume server}
+
 \end{document}

### 2012-12-14 6.858/paper/paper.tex
Spruce things up a bit

--- 2012-12-14 6.858/paper/paper.tex
@@
 authentication succeeds, the server signs the client's key and returns the
-resultant certificate.  Using these certificates, clients can then authenticate
-securely to one another without further use of the server.
+resultant certificate.  Using these certificates, clients can then prove their
+handles securely to one another without further use of the server.
+
+Note that the assigning of handles is orthogonal to the security of the protocol
+itself, and is only useful in associating human-friendly names to their actual
+cryptographic identity, i.e., their private key. Users who do not wish to use
+\Trill's federated identity system can instead self-sign their certificates and
+verify these credentials out-of-band. This is very similar to PGP's ``web of
+trust,'' and while we have not implemented such a system, we imagine the
+integration of \Trill with PGP would be trivial. An SSH-like Trust On First Use
+(TOFU) system might also be viable here.
+
+\section*{Establishing Trust}
+% Trust can be rooted two ways:
+% - Traditional CAs: Verify plume certificate against CAs, verify identities
+%   against plume certificate
+% - DNSSEC: the key (or a fingerprint of the key) for any domain can be added as
+%   a signed TXT record. We can root our trust back up to the DNSSEC root.
+% This second way seems more secure: root CAs are allowed to sign any record,
+% compromising some CAs does not seem difficult. DNS has the interesting
+% property of hierarchically namespaced trust: it is likely that a root zone
+% compromise will be difficult (root keys only used rarely) and discovered
+% quickly; owners of ".cn" keys will be unable to spoof ".com"s, etc.
+
+% Also semantics of plume CAs are "have control of this domain" => "can give out
+% identities at this domain." No communications security is based on this, only
+% naming.
 
@@
 
+\section*{Implementation}
+% trill - client library. allows plug-and-play event loops, architected to (one
+%   day!) support multiple TLS libraries.
+% plume - server. separate login and protocol servers. access to domain root
+%   keys has very small surface area. chroot more!
+
 \end{document}

### 2012-12-14 6.858/paper/paper.tex
Establishing trust

--- 2012-12-14 6.858/paper/paper.tex
@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
 
-\usepackage{fancyhdr, geometry, hyperref, sectsty, titlesec, xspace}
+\usepackage{fancyhdr, geometry, hyperref, xspace}
 \usepackage[parfill]{parskip}
@@
 
-\sectionfont{\vspace{-2pt}\large}
-\subsectionfont{\vspace{-2pt}\normalsize}
-\subsubsectionfont{\normalfont\itshape}
-
-\titlespacing\section{0pt}{10pt}{2pt}
-\titlespacing\subsection{0pt}{8pt}{0pt}
-
-\fancyhead[RO]{Carl Jackson \& Max Wang}
+\fancyhead[RO]{Carl Jackson and Max Wang}
 \fancyhead[LO]{\url{http://github.com/motmotchat/trill}}
@@
 \newcommand{\Trill}{\textsc{Trill}\xspace}
-\newcommand{\Plume}{\textsc{Plume}\xspace}
 
 \title{\Trill: Secure firewall-traversing networking}
-\author{Carl Jackson \and Max Wang}
+\author{Carl Jackson and Max Wang}
 \date{December 14, 2012}
+
+\makeatletter
+\renewcommand{\maketitle}{%
+  \thispagestyle{plain}%
+  \begin{center}%
+    {\LARGE \@title \par}%
+    {\large \@author \par}%
+    {\large \@date \par}%
+  \end{center}%
+}
+\makeatother
 
@@
 
+\section*{Overview}
 We implemented \Trill, a networking abstraction that provides secure
@@
 
-\section{Design}
-
-\subsection{Security}
+\section*{Security}
 The security of our system is based on standard cryptographic infrastructure and
@@
 
-\subsection{Firewall Traversal}
+\section*{Firewall Traversal}
 Due to restrictions on many firewalls and NATs in place on the Internet, it is
@@
 
-\section{Implementation}
+\section*{Implementation}
 % trill - client library. allows plug-and-play event loops, architected to (one
@@
 
-\subsection{Trill client library}
-
-\subsection{Plume server}
-
 \end{document}

### 2012-12-14 6.858/paper/paper.tex
Implementation section

--- 2012-12-14 6.858/paper/paper.tex
@@
 
+Our implementation of \Trill has two major components: a client library for
+connecting to peers, and a set of server implementations which handle the
+secure, federated message routing.
+
+\subsection{Plume servers}
+
+We break the server-side workload into two separate servers.  The first of
+these is the \Plume login server.  The login server bears the full
+responsibility of authenticating clients.  Each domain (or each login server on
+each domain) can determine its own method for deciding whether or not to
+endorse any given client's claimed identity.  This behavior is unspecified by
+\Trill; the login server can only endorse identities in its domain, but can
+endorse them however it sees fit.
+
+The actual process of ``logging in'' requires a client to send the login server
+a X.509 certificate signing request (CSR).  The login server, if it so chooses,
+can then endorse the client's identity by signing a cert and returning it to
+the user.  The client uses cert whenever it connects to \Plume protocol servers
+or to peers in order to assert its identity.  These servers and parties can
+then walk their way up the chain of trust to verify the user's identity and the
+authority of the login server that endorsed it.
+
+With this cert, the client is then able to connect to the \Plume protocol
+servers.  These servers will only talk over TLS to clients with valid
+identities (connections to the login servers are also secured by TLS, but only
+the login server's cert is used).  These \Plume servers server two major
+purposes:
+
+\begin{enumerate}
+  \item Message routing.  Clients can ask their \Plume servers to pass messages
+    to peers.  If they are on the same domain and are connected to the server,
+    the message is routed directly; otherwise, the message is forwarded first
+    to the proper server
+\end{enumerate}
+
+Discovery of server addresses and ports
+
 \subsection{Trill client library}
 
-\subsection{Plume server}
-
 \end{document}

### 2012-12-14 6.858/paper/paper.tex
Moar formatting

@@
 \subsection{Security}
+
 The security of our system is based on standard cryptographic infrastructure and
@@
 
-\section*{Federated Identity}
+\subsection{Federated Identity}
+
 As with other federated protocols like XMPP and email, \Trill handles are based
@@
 
-\section*{Establishing Trust}
+\subsection{Establishing Trust}
+
 Trust in \Trill is established between two clients when each of them verifies
@@
 \subsection{Firewall Traversal}
+
 Due to restrictions on many firewalls and NATs in place on the Internet, it is

### 2012-12-14 6.858/paper/paper.tex
Client mumble

@@
 \subsection{Trill client library}
+The \Trill client library is fairly straightforward. Its job is simply to open a
+UDP socket, perform a \verb`chownat`-like discovery protocol (which we call
+``probing''), and then perform a DTLS handshake. The probing protocol is the
+only non-obvious part of this. In particular, since TLS demands that one party
+act as the server and the other as the client, the probing protocol must somehow
+decide on a winner and a loser. Unfortunately, a result from the distributed
+systems literature tells us that it is impossible for two parties to both agree
+on even a single bit of information on a lossy out-of-order channel like UDP, so
+the cause seems like a lost one. However, in practice, it is possible to use
+DTLS's retry behavior, combined with non-colliding message types and
+drop-on-error packet parsing semantics, to work around this. Both parties
+generate random 64-bit numbers, and the client with the larger number is the
+winner. If a client discovers it is the winner, it transitions to the ``server''
+state, in which it attempts to read (and parse) DTLS handshake records, but
+continues to send probes to its peer at regular intervals. If a client discovers
+that it is the loser, and that its peer has acknowledged this fact (an extra bit
+in the message), it transitions to ``client'' mode, in which it attempts a DTLS
+handshake, discarding probe packets it receives. The asymmetry around the
+acknowledge bit and the timeouts allow us to guarantee protocol progress in the
+limiting case.
+
+We wanted the \Trill client library to be as flexible as possible, and a large
+portion of the implementation effort was spent making sure the library was
+separated from both the particulars of the event loop implementation and the
+choice of cryptographic library. Clients of \Trill must pass in a table of
+function pointers which allow \Trill to be notified of various event loop
+events, such as socket read/write availability and timers.  Secondly, the choice
+of cryptographic library is controllable via a preprocessor flag, and \Trill
+implementations can provide several of them in the hopes of greater
+compatibility. Although we only provided implementations for Glib and GnuTLS
+(the libraries that \verb`libpurple`, a popular instant messaging application
+uses, which was the use case we originally envisioned for \Trill), a brief
+perusal of the APIs for \verb`libev`, \verb`libevent`, OpenSSL, and NSS confirms
+that the abstractions we expose in \Trill are sufficiently general to allow any
+combination of event loop and TLS implementation to be used.
 

### 2012-12-14 6.858/paper/paper.tex
Finishing server section

@@
 Our implementation of \Trill has two major components: a client library for
-connecting to peers, and a set of server implementations which handle the
-secure, federated message routing.
+connecting to peers, and a set of server implementations (the \Plume servers)
+which handle the secure, federated message routing.
 
@@
 
-The actual process of ``logging in'' requires a client to send the login server
-a X.509 certificate signing request (CSR).  The login server, if it so chooses,
-can then endorse the client's identity by signing a cert and returning it to
-the user.  The client uses cert whenever it connects to \Plume protocol servers
-or to peers in order to assert its identity.  These servers and parties can
-then walk their way up the chain of trust to verify the user's identity and the
-authority of the login server that endorsed it.
+The actual process of ``logging in'' requires a client to connect to the login
+server---over TLS-secured TCP using the login server's X.509 cert---send the
+login server a certificate signing request (CSR).  The login server, if it so
+chooses, can endorse the client's identity by signing a cert and returning it
+to the client.  The client then uses cert whenever it connects to \Plume
+protocol servers or to peers in order to assert its identity.  These servers
+and parties can then walk their way up the chain of trust to verify the
+client's identity and the authority of the login server that endorsed it.
 
 With this cert, the client is then able to connect to the \Plume protocol
-servers.  These servers will only talk over TLS to clients with valid
-identities (connections to the login servers are also secured by TLS, but only
-the login server's cert is used).  These \Plume servers server two major
-purposes:
+servers (or simply, \Plume servers).  These servers will only talk over TLS to
+clients with valid identities.  These \Plume servers server two major purposes:
 
 \begin{enumerate}
-  \item Message routing.  Clients can ask their \Plume servers to pass messages
-    to peers.  If they are on the same domain and are connected to the server,
-    the message is routed directly; otherwise, the message is forwarded first
-    to the proper server
+  \item \textbf{Message routing.}  Clients can ask their \Plume servers to pass
+    messages to peers.  If they are on the same domain and are connected to the
+    server, the message is routed directly; otherwise, the message is forwarded
+    first to the proper server.  Messages generally take the format of an
+    opstring, the peer's handle, the sender's cert, and a payload.
+
+  \item \textbf{UDP reflection.}  Before connecting to a peer, a client can
+    open up an ephemeral UDP port for listening and ask its \Plume server to
+    identify its IP and port, since firewalls may prevent the client from
+    finding out on its own.  In this protocol, the client first sends the
+    server a random 64-bit number alongside a reflection request.  The server
+    hashes this number with the client's handle and sends the client an ack.
+
+    At this point, the client begins regularly pinging the server with UDP
+    datagrams containing the random number.  When the server receives a UDP
+    packet that matches a client who has requested UDP reflection, it
+    determines the IP and port the client was sending on and returns it, over
+    TCP, back to the client, who saves it and halts UDP transmission.
 \end{enumerate}
 
-Discovery of server addresses and ports
+There are two major types of \Trill messages, corresponding to requests a
+client makes of its peer, that \Plume servers are responsible for routing:
+
+\begin{enumerate}
+  \item \textbf{Connect.}  To connect to a peer, a client first prepares a
+    payload used to identify itself---we colloquially refer to this as a
+    ``calling card.''  This calling card consists of the IP and port of the
+    client's UDP listener, signed by the client's identity certificate, and
+    then encrypted by the peer's public key.  Signing the cert enables the peer
+    to confirm the association between the client and the given IP and port.
+    Encryption lets the client keep its IP and port secret from other domains'
+    \Plume servers, which it may not trust.  (As clients, we necessarily have
+    to trust other domains to make correct endorsements of identity, but we can
+    also verify these endorsements out-of-band with our peers).
+
+    Encryption also prevents replay attacks in the event that someone who isn't
+    our peer obtains our calling card.  Passing just the signed IP and port
+    would allow an adversary to forge connection requests in our name, which
+    could flood us with connection responses.  Encryption makes it so that only
+    our peer can do so.
+
+  \item \textbf{Identify.}  If a client wishes to connect to a peer, it needs
+    to obtain the peer's public key in order to encrypt a calling card.  In
+    addition to connection requests, client's can also make identification
+    requests of their peers.  Willing peers can simply pass their cert back to
+    the requesting client, who can then initiate connection.
+\end{enumerate}
+
+Discovery of server addresses and ports is done via DNS queries for SRV
+records; we specify the use of \verb`_plume` and \verb`_plume-login`, both over
+TCP, for the \Plume and login servers, respectively, and \verb`_plume-udp` for
+the UDP reflection service.  All messages sent over the wire are marshalled
+using \textsc{MessagePack}.
+
+Once two peers have exchanged calling cards, they can then use the \Trill
+library to establish a secured UDP connection.
 

### 2012-12-26 src/plume.c
Fixing Plume client and mkcrt.bash
Note that the Plume client requires the glib-networking package for TLS
to work properly.

@@ -13,8 +13,6 @@ int
 plume_tls_verify_cert(GTlsConnection *conn, GTlsCertificate *cert,
     GTlsCertificateFlags errors, void *data)
 {
-  log_warn("plume_tls_verify_cert");
-
   return 1;
 }
 
@@ -25,7 +23,9 @@ plume_tls_setup(GSocketClient *client, GSocketClientEvent event,
   GTlsCertificate *cert;
   GTlsClientConnection *tls_conn;
 
-  log_warn("plume_tls_setup: %d", event);
+  if (event == G_SOCKET_CLIENT_TLS_HANDSHAKED) {
+    log_warn("TLS handshake completed");
+  }
 
   if (event != G_SOCKET_CLIENT_TLS_HANDSHAKING) {
     return;
@@ -33,8 +33,11 @@ plume_tls_setup(GSocketClient *client, GSocketClientEvent event,
 
   tls_conn = (GTlsClientConnection *)conn;
 
-  cert = g_tls_certificate_new_from_file("client.crt", NULL);
+  // Bind our certificate to the connection.
+  cert = g_tls_certificate_new_from_files("client.crt", "client.key", NULL);
   g_tls_connection_set_certificate((GTlsConnection *)tls_conn, cert);
+
+  // Validate ALL the server certs!
   g_tls_client_connection_set_validation_flags(tls_conn, 0);
 
   g_signal_connect(tls_conn, "accept-certificate",
@@ -54,6 +57,7 @@ plume_connect_server(GObject *obj, GAsyncResult *res, void *data)
   resolver = (GResolver *)obj;
   srv = g_resolver_lookup_service_finish(resolver, res, NULL)->data;
 
+  // Construct a client socket for the service.
   client = g_socket_client_new();
   g_socket_client_set_tls(client, TRUE);
   addr = g_network_address_new(g_srv_target_get_hostname(srv),
@@ -69,6 +73,7 @@ main(int argc, char *argv[])
 {
   g_type_init();
 
+  // Look up the Plume server.
   g_resolver_lookup_service_async(g_resolver_get_default(), "plume", "tcp",
       "mxawng.com", NULL, plume_connect_server, NULL);
 

### 2012-12-26 src/plume.c
Abort Plume client if TLS is unsupported

@@ -73,6 +73,13 @@ main(int argc, char *argv[])
 {
   g_type_init();
 
+  // Abort if Glib doesn't have a real TLS backend.
+  if (!g_tls_backend_supports_tls(g_tls_backend_get_default())) {
+    log_error("No TLS support found; try installing the "
+              "glib-networking package.");
+    return 1;
+  }
+
   // Look up the Plume server.
   g_resolver_lookup_service_async(g_resolver_get_default(), "plume", "tcp",
       "mxawng.com", NULL, plume_connect_server, NULL);
